---
title: "Power Analysis"
author: "Mac Prible"
date: "2024-03-30"
output: html_document
---


https://pubmed.ncbi.nlm.nih.gov/22491174/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("power_analysis_simulation.R")
```

# Study 2
## Hypothesis: SLR moves significantly in same direction in all conditions

Planned comparisons are made between baseline and post adaptation within each condition. 
```{r}

# healthy population...anticipate symmetry
pre_means <-list(
  "cSBT" = 1,
  "FastProp" = 1,
  "FastBrake" = 1
)

post_means <-list(
  "cSBT" = 1.49,
  "FastProp" = 1.26, 
  "FastBrake" = 1.31 
)


std_dev <- .045
pre_std_devs <-list(
  "cSBT" = std_dev,
  "FastProp" = std_dev,
  "FastBrake" = std_dev
)

std_dev <- 0.32
post_std_devs <-list(
  "cSBT" = std_dev,
  "FastProp" = std_dev,
  "FastBrake" = std_dev
)
    
conditions <- list( "cSBT", "FastProp", "FastBrake" )
    
test_sample_sizes = 10:15
for (n in test_sample_sizes){
  
    num_simulations = 20
    
    # Running simulations in parallel
    aim_2_slr_results <- lapply(1:num_simulations, function(i) {
      simulate_aim(n, conditions, pre_means, pre_std_devs, post_means, post_std_devs, correlation = 0.5, FastPropOnly = FALSE)
    })
    
    # Computing success count and success rate
    success_count <- sum(unlist(aim_2_slr_results))
    success_rate <- success_count / num_simulations
    
    print(paste("For n = ", n, ", success rate is ", success_rate, " after ", num_simulations, " simulations."))
  
}

```


## Hypothesis: PF impulse increases in FastProp, decreases in other conditions

Powered to distinguish differences in interaction effect (i.e. PDSV may have half impact on SLR)
```{r}

baseline_mean <- 0.415
pre_means <-list(
  "cSBT" = baseline_mean,
  "FastProp" = baseline_mean,
  "FastBrake" =baseline_mean
)

post_means <-list(
  "cSBT" = .32,
  "FastProp" = .44, 
  "FastBrake" = .30
)

std_dev <- .023
pre_std_devs <-list(
  "cSBT" = std_dev,
  "FastProp" = std_dev,
  "FastBrake" = std_dev
)

std_dev <- .068
post_std_devs <-list(
  "cSBT" = std_dev,
  "FastProp" = std_dev,
  "FastBrake" = std_dev
)

conditions <- list( "cSBT", "FastProp", "FastBrake" )
       
test_sample_sizes = 12:16
for (n in test_sample_sizes){
  
    num_simulations = 20
    
    # Running simulations in parallel
    aim_2_pf_results <- lapply(1:num_simulations, function(i) {
      simulate_aim(n, conditions, pre_means, pre_std_devs, post_means, post_std_devs, correlation=0.5, FastPropOnly = TRUE)
    })
    
    # Computing success count and success rate
    success_count <- sum(unlist(aim_2_pf_results))
    success_rate <- success_count / num_simulations
    
    print(paste("For n = ", n, ", success rate is ", success_rate, " after ", num_simulations, " simulations."))
  
}

```


# Study 3

## Change PF Impulse Bilaterally in either direction


```{r}

# estimates here taken from subject 1 values; time normalized impulse across stance
pre_means <-list(
  "FastProp" = .40,
  "FastBrake" = .40
)

# subject 1: PF impulse decrease in cSBT (as with Lauziere 2014) and fastBrake, increased with FastProp
post_means <-list(
  "FastProp" = 0.50,
  "FastBrake" = 0.36
)

# estimates based on preliminary data (5 subjects)
pre_vars <-list(
  "FastProp" = .032,
  "FastBrake" = .032
)

# expecting increased variance following perturbation
post_vars <-list(
  "FastProp" = .087,
  "FastBrake" = .087
)
    
conditions <- list( "FastProp", "FastBrake" )
    
test_sample_sizes = 5:8
for (n in test_sample_sizes){
  
    num_simulations = 20
    
    # Running simulations in parallel
    aim_3_pf_results <- lapply(1:num_simulations, function(i) {
      simulate_aim(n, conditions, pre_means, pre_vars, post_means, post_vars, correlation = 0.0, FastPropOnly = TRUE)
    })
    
    # Computing success count and success rate
    success_count <- sum(unlist(aim_3_pf_results))
    success_rate <- success_count / num_simulations
    
    print(paste("For n = ", n, ", success rate is ", success_rate, " after ", num_simulations, " simulations."))
  
}

```