---
title: "Primary Analysis"
author: "Mac Prible"
output: html_document
---

# Import Data


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
options(error = function() traceback(2))
```


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(emmeans)

tidy_data_directory <- "C:\\Users\\Mac Prible\\OneDrive - The University of Texas at Austin\\research\\PDSV\\data\\dissertation\\v3d\\tidy_output"

figure_directory <- "C:\\Users\\Mac Prible\\OneDrive - The University of Texas at Austin\\research\\PDSV\\data\\dissertation\\figures"

mean_across_stance <- 
  list.files(tidy_data_directory, pattern="*_mean_across_stance.csv") %>% 
  map_df(~read_csv2(file.path(tidy_data_directory,.)))
  
step_lengths <- 
  list.files(tidy_data_directory, pattern="*_step_lengths.csv") %>% 
  map_df(~read_csv2(file.path(tidy_data_directory,.)))
```

## Get Stance Times

```{r}
stance_times <- mean_across_stance %>% 
  filter(variable %in% c("TIME") & axis == "X") %>% 
  group_by(subject, condition, period, stance_side, normalized_time) %>% 
  summarize(time = mean(value)) %>% 
  ungroup() %>% 
  group_by(subject, condition, period, stance_side) %>%   
  summarize(start_stance = min(time),
            stop_stance = max(time)) %>% 
  ungroup() %>% 
  mutate(stance_time = stop_stance - start_stance)

```

## Define Conditions

```{r}
UNI_CONDITION_LABELS <- c(
  "sbt" = "Conventional SBT",
  "ubp" = "Uni. Fast Brake",
  "upp" = "Uni. Fast Prop"
)

BIL_CONDITION_LABELS <- c(
  "bbp" = "Bil. Fast Brake",
  "bpp" = "Bil. Fast Prop"
)

CONDITION_LABELS <- c(UNI_CONDITION_LABELS, BIL_CONDITION_LABELS)
```

# Belt Speed

```{r}


mean_across_stance %>% 
  filter(subject=="s1") %>% 
  filter(period == "Late Adapt") %>% 
  filter(condition == "sbt") %>% 
  mutate(condition = CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_BeltSpeed") & axis == "X") %>%
  # filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  ggplot(aes(x=normalized_time,y= value, linetype=period, color=stance_side))+
  geom_line(size=1.2)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(subject~condition + stance_side)+
  theme_minimal()+
  theme(legend.position="none")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Subject 1: Actual Belt Speed for Conventional Split Belt Treadmill",
    subtitle = "Averaged Across 5 Steps",
    color = "Condition"
  )+
  scale_color_manual(values= c("left" = "blue", "right" = "red"))  # Added color specification


ggsave(file.path(figure_directory,"Conventional SBT speeds.png"))

mean_across_stance %>% 
  filter(subject=="s1") %>% 
  filter(period == "Late Adapt") %>% 
  filter(condition == "ubp") %>% 
  mutate(condition = CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_BeltSpeed") & axis == "X") %>%
  # filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  ggplot(aes(x=normalized_time,y= value, linetype=period, color=stance_side))+
  geom_line(size=1.2)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(subject~condition + stance_side)+
  theme_minimal()+
  theme(legend.position="none")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Subject 1: Actual Belt Speed for Unilateral FastBrake",
    subtitle = "Averaged Across 5 Steps",
    color = "Condition"
  )+
  scale_color_manual(values= c("left" = "blue", "right" = "red"))  # Added color specification


ggsave(file.path(figure_directory,"Uni fast brake speeds.png"))

mean_across_stance %>% 
  filter(subject=="s1") %>% 
  filter(period == "Late Adapt") %>% 
  filter(condition == "upp") %>% 
  mutate(condition = CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_BeltSpeed") & axis == "X") %>%
  # filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  ggplot(aes(x=normalized_time,y= value, linetype=period, color=stance_side))+
  geom_line(size=1.2)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(subject~condition + stance_side)+
  theme_minimal()+
  theme(legend.position="none")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Subject 1: Actual Belt Speed for Unilateral FastProp",
    subtitle = "Averaged Across 5 Steps",
    color = "Condition"
  )+
  scale_color_manual(values= c("left" = "blue", "right" = "red"))  # Added color specification


ggsave(file.path(figure_directory,"Uni fast prop speeds.png"))
```

```{r}

mean_across_stance %>% 
  filter(subject=="s1") %>% 
  filter(period == "Late Adapt") %>% 
  filter(condition %in% c("upp","ubp", "sbt")) %>% 
  mutate(condition = CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_BeltSpeed") & axis == "X") %>%
  # filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  ggplot(aes(x=normalized_time,y= value, linetype=period, color=stance_side))+
  geom_line(size=1.2)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(rows=vars(condition), cols = vars(stance_side))+
  theme_minimal()+
  theme(legend.position="none")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Subject 1: Actual Belt Speed Across All Unilateral Conditions",
    subtitle = "Averaged Across 5 Steps",
    color = "Condition"
  )+
  scale_color_manual(values= c("left" = "blue", "right" = "red"))  # Added color specification


ggsave(file.path(figure_directory,"All Uni Condition speeds.png"))

mean_across_stance %>% 
  filter(subject=="s1") %>% 
  filter(period == "Late Adapt") %>% 
  filter(condition %in% c("bpp","bbp")) %>% 
  mutate(condition = CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_BeltSpeed") & axis == "X") %>%
  # filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  ggplot(aes(x=normalized_time,y= value, linetype=period, color=stance_side))+
  geom_line(size=1.2)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(rows=vars(condition), cols = vars(stance_side))+
  theme_minimal()+
  theme(legend.position="none")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Subject 1: Actual Belt Speed Across All Bilateral Conditions",
    subtitle = "Averaged Across 5 Steps",
    color = "Condition"
  )+
  scale_color_manual(values= c("left" = "blue", "right" = "red"))  # Added color specification


ggsave(file.path(figure_directory,"All Bilateral Condition speeds.png"))
```

# Step Length Ratio (SLR)

## All time periods.

```{r}
primary_periods <- c("Late Baseline", "Early Adapt" , "Late Adapt", "Early Post Adapt")
 
plt_SLR <- step_lengths %>% 
  filter(period %in% primary_periods) %>% 
  filter(condition %in% names(UNI_CONDITION_LABELS)) %>% 
  mutate(condition = UNI_CONDITION_LABELS[condition]) %>% 
  mutate(period = factor(period, levels = primary_periods)) %>% 
  group_by(subject, condition, period, stance_side) %>% 
  summarize(step_length = mean(step_length)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = stance_side, values_from = step_length, names_glue = "{stance_side}_step_length") %>% 
  mutate(SLR = left_step_length/right_step_length) %>% 
  ggplot(aes(x=period, y = SLR, group = subject, color=subject))+
  geom_point()+
  geom_line()+
  geom_hline(yintercept =1, linetype="dashed")+
  facet_grid(cols=vars(condition))+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5,),
        plot.caption = element_text(hjust = 0.5, size = 13),
        axis.text.x = element_text(angle = 45, hjust = 1),   # Add this line
        legend.position = "none")+
  labs(title = "SLR Across Treadmill Protocol")

# plt_SLR
# 
# ggsave(file.path(figure_directory,"SLR_all.png"))

plt_SLR +
  stat_summary(aes(group = 1), fun = mean, geom = "line", color = "black", size = 0.8, linetype = "dashed", alpha=.6) +  # Thinner, dashed mean line
  stat_summary(aes(group = 1), fun = mean, geom = "point", color = "black", size = 3, shape = 18) +  # Smaller mean points
  geom_text(stat = "summary", fun = mean,
            aes(label = sprintf("%.2f", ..y..), group = 1,
                # hjust = ifelse(period %in% c("Baseline"), 1.5, -1)
                ),
            color = "black", vjust = -0.5, size = 4, fontface = "bold",
            bg.color = "white", bg.r = 0.15)

ggsave(file.path(figure_directory,"SLR_all_w_mean.png"))

```

## Late Baseline vs Early Post Adaptation

```{r}

pre_post_periods <- c("Late Baseline", "Early Post Adapt")


SLR_data <- step_lengths %>% 
  filter(period %in% pre_post_periods) %>% 
  filter(condition %in% names(UNI_CONDITION_LABELS)) %>% 
  mutate(condition = CONDITION_LABELS[condition]) %>% 
  mutate(period = factor(period, levels = pre_post_periods)) %>% 
  group_by(subject, condition, period, stance_side) %>% 
  summarize(step_length = mean(step_length)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = stance_side, values_from = step_length, names_glue = "{stance_side}_step_length") %>% 
  mutate(SLR = left_step_length/right_step_length)
 
plt_SLR <-  SLR_data %>% 
  ggplot(aes(x=period, y = SLR, group = subject, color=subject))+
  geom_point()+
  geom_line()+
  geom_hline(yintercept =1, linetype="dashed")+
  facet_grid(cols=vars(condition))+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5,),
        plot.caption = element_text(hjust = 0.5, size = 13),
        axis.text.x = element_text(angle = 45, hjust = 1),   # Add this line
        legend.position = "none")+
  labs(title = "SLR Across Treadmill Protocol")

# plt_SLR
# 
# ggsave(file.path(figure_directory,"SLR_pre_post.png"))

plt_SLR +
  stat_summary(aes(group = 1), fun = mean, geom = "line", color = "black", size = 0.8, linetype = "dashed", alpha=.6) +  # Thinner, dashed mean line
  stat_summary(aes(group = 1), fun = mean, geom = "point", color = "black", size = 3, shape = 18) +  # Smaller mean points
  geom_text(stat = "summary", fun = mean,
            aes(label = sprintf("%.2f", ..y..), group = 1,
                # hjust = ifelse(period %in% c("Baseline"), 1.5, -1)
                ),
            color = "black", vjust = -0.5, size = 4, fontface = "bold",
            bg.color = "white", bg.r = 0.15)

ggsave(file.path(figure_directory,"SLR_pre_post_mean.png"))

```
## Test

```{r}
# Prepare data for lmer
SLR_data$subject <- factor(SLR_data$subject)
SLR_data$condition <- factor(SLR_data$condition)
SLR_data$period <- factor(SLR_data$period)

model <- lmer(SLR ~ period * condition + (1|subject), 
             data = SLR_data)


summary(model)

# Compute estimated marginal means
emm <- emmeans(model, specs = ~ period | condition)

# Pairwise comparisons of Period within each Condition
period_comparisons <- pairs(emm, simple = "period", adjust="holm")
summary(period_comparisons)
```

# Ankle PF Torque

## Why is normalized time getting edge cropped?

Looks like there are NAs getting propagated. These are in the mean stance values so I need to go back to the step-by-step values to see where the NAs are coming from.
```{r}


# mean_across_stance %>% 
#   filter(period %in% primary_periods) %>% 
#   mutate(period = factor(period,levels = primary_periods)) %>% 
#   filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
#   filter(is.na(value)) %>% view()
#   ungroup() %>% 
#   
#   group_by(order, condition, stance_side, variable, axis,normalized_time) %>% 
#   summarize(value = mean(value)) %>% 
#   view()
```

## Unilateral PDSV

### Characteristic Torque Profile

```{r, fig.height=10, fig.width=8}

mean_across_stance %>% 
  filter(period %in% pre_post_periods) %>% 
  mutate(period = factor(period,levels = pre_post_periods)) %>% 
  filter(condition %in% names(UNI_CONDITION_LABELS)) %>%  
  mutate(condition = UNI_CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  group_by(period, condition, variable, normalized_time, stance_side) %>% 
  summarize(
    PF_torque = mean(value),
    se = sd(value)/sqrt(n()),  # Standard error
    ci = 1.96 * se,            # 95% confidence interval
    upper = PF_torque + ci,
    lower = PF_torque - ci,
    .groups = 'drop'
  ) %>% 
  ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
  geom_ribbon(aes(
    ymin = lower, 
    ymax = upper,
    fill = condition,
    group = interaction(condition, period)
  ), 
  alpha = 0.2,
  color = NA
  ) +
  geom_line(aes(y = PF_torque), size = 1) +
  geom_hline(yintercept = 0) +
  scale_x_continuous(labels = NULL) +
  facet_grid(
    rows = vars(condition),
    cols = vars(stance_side)
  ) +
  scale_linetype_manual(values = c("Early Post Adapt" = "solid", "Late Baseline" = "dashed")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Ankle PF Torque (N.m/BW)",
    title = "Ankle Torque Across Conditions",
    subtitle = "Averaged Across Subjects (95% CI)",
    color = "Condition",
    fill = "Condition"
  )


ggsave(file.path(figure_directory,"unil_torque_over_time.png"))
```

### PF Impulse

```{r calculate_impulse}


PF_impulse <- mean_across_stance %>% 
  filter(period %in% pre_post_periods) %>% 
  mutate(period = factor(period,levels = pre_post_periods)) %>% 
  filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  group_by(subject, period, condition, variable, normalized_time,stance_side) %>% 
  summarize(PF_torque = mean(value)) %>% 
  filter(PF_torque<0) %>%
  group_by(subject, stance_side,condition, period) %>% 
  summarize(PF_impulse = sum(PF_torque)/-100)


```

#### Plot
```{r, fig.width=6, fig.height=6}

plt <- PF_impulse %>% 
  ungroup() %>% 
  filter(condition %in% names(UNI_CONDITION_LABELS)) %>%  
  mutate(condition = UNI_CONDITION_LABELS[condition]) %>% 
  filter(stance_side=="left") %>% 
  ggplot(aes(x = period,y=PF_impulse, group = subject, color=subject))+
  geom_point()+
  geom_line()+
  facet_grid(~condition )+
  # scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")+
  labs(
    x = "Period",
    y = "Ankle PF Impulse (N.m.s/BW)",
    title = "Ankle PF Impulse Across Periods",
  )

# plt 
# ggsave(file.path(figure_directory,"uni_impulse.png"))

plt+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6) +
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) +
  # Optionally add mean values as text
  geom_text(stat = "summary", 
            fun = mean,
            aes(label = sprintf("%.2f", ..y..), 
                hjust = ifelse(period=="Late Baseline", 1.5, -0.5),
                group = 1),
            color = "black", 
            vjust = -0.5, 
          
            size = 3)

ggsave(file.path(figure_directory,"uni_impulse_mean.png"))
```

#### Test

```{r}


unilateral_data <- PF_impulse %>% 
  ungroup() %>% 
  filter(condition %in% names(UNI_CONDITION_LABELS)) %>%  
  mutate(condition = UNI_CONDITION_LABELS[condition]) %>% 
  filter(stance_side=="left")
  
# Prepare data for lmer
unilateral_data$subject <- factor(unilateral_data$subject)
unilateral_data$condition <- factor(unilateral_data$condition)
unilateral_data$period <- factor(unilateral_data$period)

model <- lmer(PF_impulse ~ period * condition + (1|subject), 
             data = unilateral_data)


summary(model)

# Compute estimated marginal means
emm <- emmeans(model, specs = ~ period | condition)

# Pairwise comparisons of Period within each Condition
period_comparisons <- pairs(emm, simple = "period", adjust="holm")
summary(period_comparisons)
```


## Bilateral PDSV

### Characteristic Torque Profile
```{r, fig.height=7, fig.width=8}

mean_across_stance %>% 
  filter(period %in% pre_post_periods) %>% 
  mutate(period = factor(period,levels = pre_post_periods)) %>% 
  filter(condition %in% names(BIL_CONDITION_LABELS)) %>%  
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  group_by(period, condition, variable, normalized_time, stance_side) %>% 
  summarize(
    PF_torque = mean(value),
    se = sd(value)/sqrt(n()),  # Standard error
    ci = 1.96 * se,            # 95% confidence interval
    upper = PF_torque + ci,
    lower = PF_torque - ci,
    .groups = 'drop'
  ) %>% 
  ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
  geom_ribbon(aes(
    ymin = lower, 
    ymax = upper,
    fill = condition,
    group = interaction(condition, period)
  ), 
  alpha = 0.2,
  color = NA
  ) +
  geom_line(aes(y = PF_torque), size = 1) +
  geom_hline(yintercept = 0) +
  scale_x_continuous(labels = NULL) +
  facet_grid(
    rows = vars(condition),
    cols = vars(stance_side)
  ) +
  scale_linetype_manual(values = c("Early Post Adapt" = "solid", "Late Baseline" = "dashed")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Ankle PF Torque (N.m/BW)",
    title = "Ankle Torque Across Conditions",
    subtitle = "Averaged Across Subjects (95% CI)",
    color = "Condition",
    fill = "Condition"
  )


ggsave(file.path(figure_directory,"bil_torque_over_time.png"))
```

### PF Impulse
#### Plot
```{r, fig.width=6, fig.height=6}
plt <- PF_impulse %>% 
  ungroup() %>% 
  filter(condition %in% names(BIL_CONDITION_LABELS)) %>%  
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  group_by(subject,condition, period) %>% 
  summarize(PF_impulse= mean(PF_impulse)) %>%  
  ggplot(aes(x = period,y=PF_impulse, group = subject, color=subject))+
  geom_point()+
  geom_line()+
  facet_grid( ~ condition)+
  # scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")+
  labs(
    x = "Period",
    y = "Ankle PF Impulse (N.m.s/BW)",
    title = "Ankle PF Impulse Across Periods",
  )

# plt
# 
# ggsave(file.path(figure_directory,"bil_impulse.png"))

plt+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6) +
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) +
  # Optionally add mean values as text
  geom_text(stat = "summary", 
            fun = mean,
            aes(label = sprintf("%.2f", ..y..), 
                hjust = ifelse(period=="Late Baseline", 1.5, -0.5),
                group = 1),
            color = "black", 
            vjust = -0.5, 
          
            size = 3)

ggsave(file.path(figure_directory,"bil_impulse_w_mean.png"))
```

#### Test

```{r}
bilateral_data <- PF_impulse %>% 
  ungroup() %>% 
  filter(condition %in% names(BIL_CONDITION_LABELS)) %>%  
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  group_by(subject,condition, period) %>%
  summarize(PF_impulse= mean(PF_impulse))
  
# Prepare data for lmer
bilateral_data$subject <- factor(bilateral_data$subject)
bilateral_data$condition <- factor(bilateral_data$condition)
bilateral_data$period <- factor(bilateral_data$period)

model <- lmer(PF_impulse ~ period * condition + (1|subject), 
             data = bilateral_data)


summary(model)

# Compute estimated marginal means
emm <- emmeans(model, specs = ~ period | condition)

# Pairwise comparisons of Period within each Condition
period_comparisons <- pairs(emm, simple = "period", adjust="holm")
summary(period_comparisons)
```


# Addendum
Just some random things I am looking at.

## Pelvis Height
```{r, fig.width=10, fig.height=6}
periods = c("Late Baseline", "Early Post Adapt")

mean_across_stance %>% 
  filter(period %in% periods) %>% 
  mutate(period = factor(period,levels = periods)) %>% 
  filter(variable %in% c("PELVIS_COM") & axis == "Y") %>% 
  group_by(period, condition, variable, normalized_time,stance_side) %>% 
  summarize(PF_torque = mean(value)) %>% 
  ggplot(aes(x=normalized_time,y= PF_torque, linetype=period, color=condition))+
  geom_line()+
  scale_x_continuous(labels=NULL)+
  facet_grid(rows=vars(condition),
             cols=vars(stance_side))+
  scale_linetype_manual(values=c("Early Post Adapt"="solid", "Late Baseline"="dashed")) +
  # theme(legend.position="bottom")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Pelvis COM AP Position",
    title = "Mean Across 5 Stance Phases",
    subtitle = "Averaged Across Subjects",
    color = "Condition"
  )
```
