---
title: "Processing V3D Data in R"
author: "Mac Prible"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("v3d_to_tidy.R")
```

# Reproducible Research

You want to be able to recreate your entire pipeline of data processing from raw data to statistical results with minimial one-off adjustments.

Avoid doing things by hand with the idea that you'll do it once and be done with it. 
You will always have to go back to revise.
Hand edits will be forgotten.
Automate everything and make it explicit.

You want to be able to rerun the processing pipeline with your original data and get the same results.
If you do things by hand or in excel, then you cannot do that.
You also cannot scale it. 

# Clean up Header to Long Metadata
```{r, warning=FALSE}
# Load required libraries
library(tidyverse)

# Set up variables
data_directory <- "C:\\Users\\Mac Prible\\OneDrive - The University of Texas at Austin\\research\\PDSV\\data\\PDVS_2024\\v3d"
file_name <- "left_v3d_in_R_gait_cycle_data.tsv"
stance_side <- "left"

input_path <- file.path(data_directory, file_name)

#tidy_data <- tsv_to_tidy(data_directory, file_name, stance_side="left")

############# Stage 1: Organize Metadata Header and get step count
# Step 1: Read the metadata
metadata <- read_tsv(input_path, n_max = 5, col_names = FALSE, show_col_types = FALSE)

view(metadata)
# Step 2: Transpose the metadata
metadata <- metadata %>%
  t() %>%
  as_tibble(.name_repair = "unique") %>%
  setNames(c("filename", "raw_variable", "type", "origin", "axis")) %>% 
  select(-type, -origin) # cannot see a use for these

view(metadata)

# Step 3: Track column order and clean up cruff (can't have sorted data up to this point)
metadata <- metadata %>%
  mutate(column_number = row_number()) %>% 
  slice(-1) # first column (now row) is just NA

view(metadata)

# Step 4: Clean up variable names to make side reflect IPSI or CONTRA in terms of stance side frame of reference.
metadata <- metadata %>% 
  mutate(stance_side = stance_side) %>% 
  mutate(variable = case_when(
         raw_variable == "RHEEL" ~ "R_HEEL",
         raw_variable == "LHEEL" ~ "L_HEEL",
         raw_variable == "FP1" ~ "Left_GRF",
         raw_variable == "FP2" ~ "Right_GRF",
         TRUE ~ raw_variable  # This keeps all other values unchanged
       ))  %>% 
  mutate(variable = case_when(
    stance_side == "left" & grepl("^(L_|Left)", variable) ~ sub("^(L_|Left_|Left)", "IPSI_", variable),
    stance_side == "right" & grepl("^(R_|Right)", variable) ~ sub("^(R_|Right_|Right)", "IPSI_", variable),
    stance_side == "right" & grepl("^(L_|Left)", variable) ~ sub("^(L_|Left_|Left)", "CONTRA_", variable),
    stance_side == "left" & grepl("^(R_|Right)", variable) ~ sub("^(R_|Right_|Right)", "CONTRA_", variable),
    TRUE ~ variable # Keeps all other values unchanged
  ))  

view(metadata)

# Step 5: Sort by combined header data and column number, then assign step count
metadata <- metadata %>%
  mutate(combined_header_data = paste0(filename, variable, axis, stance_side)) %>% 
  arrange(combined_header_data, column_number) %>%
  group_by(combined_header_data) %>%
  mutate(step = row_number()) %>%
  ungroup() %>% 
  select(-combined_header_data)

view(metadata)

# Step 6: unpack filename into meaningful columns
metadata <- metadata %>% 
  separate(col=filename, into = c("subject", "order", "condition", "start_stop"), sep = "_") %>%
  mutate(start_stop = str_replace(start_stop, ".c3d","")) 

view(metadata)

# Step 7: Map to Raw Data Column Names
metadata <- metadata %>%
  select(subject, order, condition, start_stop, stance_side, variable, axis, column_number, step) %>%
  mutate(column_id = paste0("X", column_number)) 

view(metadata)

```


# Fold Raw Data into Long Metadata

```{r}

# 1. Read the raw data
raw_data <- read_tsv(input_path, 
                     col_names = FALSE, 
                     skip = 5, 
                     show_col_types = FALSE,
                     col_types = cols(.default = col_double()))
view(raw_data)
# 2. Reshape the raw data
long_data <- raw_data %>%
  rename(normalized_time = X1) %>% 
  pivot_longer(cols = -normalized_time, names_to = "column_id", values_to = "value") 

view(long_data)

# 3. Join the metadata with the reshaped raw data
combined_data <- long_data %>%
  left_join(metadata, by = "column_id")

view(combined_data)

# 4. Clean up and arrange the final dataset
tidy_data <- combined_data %>%
  select(subject, order, condition, start_stop, stance_side, variable,  axis, step, normalized_time, value) 

```



# Find the steps to put into each period

## examine the belt speeds
```{r}

belt_speeds <- tidy_data %>% 
  ungroup() %>% 
  filter(str_detect(variable, "Belt")) %>% 
  mutate(value = 1000*value) %>% # correct scale
  pivot_wider(names_from = variable, values_from = value) %>% 
  mutate(speed_difference = IPSI_BeltSpeed- CONTRA_BeltSpeed,
         abs_speed_difference = abs(speed_difference))

  
belt_speeds %>% 
  ggplot(aes(x = normalized_time, y = IPSI_BeltSpeed, group=step))+
  geom_line()+
  facet_grid(rows=vars(condition), cols = vars(start_stop))

```

## Get max speed difference per step
```{r}

max_speed_diff <- belt_speeds %>% 
  group_by(subject, order,condition,start_stop, stance_side,axis, step) %>% 
  summarize(max_speed_diff = max(abs_speed_difference)) 

max_speed_diff %>%  
  ggplot(aes(x = step, y = max_speed_diff))+
  geom_point()+
  facet_grid(rows=vars(condition), cols = vars(start_stop))

```

## Find when change occurs
```{r}


find_change_point <- function(data, window_size = 5) {
  n <- nrow(data)
  max_diff <- 0 # this will get updated as 
  change_point <- NA
  
  for (i in (window_size + 1):(n - window_size)) {
    mean1 <- mean(data$max_speed_diff[(i-window_size):(i-1)])
    mean2 <- mean(data$max_speed_diff[i:(i+window_size-1)])
    diff <- abs(mean2 - mean1)
    if (diff > max_diff) {
      max_diff <- diff
      change_point <- data$step[i]
    }
  }
  
  return(change_point)
}

speed_change_steps <- max_speed_diff %>%
  group_by(condition, start_stop) %>%
  summarize(change_point = find_change_point(cur_data())) %>% 
  ungroup()

# Add the change points to the original plot
steps_by_period %>%
  ggplot(aes(x = step, y = max_speed_diff)) +
  geom_point() +
  geom_vline(data = speed_change_steps, aes(xintercept = change_point), color = "red", linetype = "dashed") +
  facet_grid(rows = vars(condition), cols = vars(start_stop)) +
  labs(title = "Belt Speed Difference with Detected Change Points",
       x = "Step",
       y = "Max Speed Difference")+
  theme(legend.position = "bottom") 


```


## Assign Period based on Speed Change
```{r}

steps_to_include <- 5
# Join the change points with the original data
steps_by_period <- max_speed_diff %>%
  ungroup() %>% 
  select(-axis) %>%  
  left_join(speed_change_steps, by = c("condition", "start_stop")) %>%
  # group_by(condition, start_stop) %>%
  mutate(period = case_when(
    start_stop == "start" & step >= (change_point - steps_to_include -1) & step < change_point-1 ~ "Late Baseline",
    start_stop == "start" & step > change_point & step < (change_point + steps_to_include +1 ) ~ "Early Adaptation",
    start_stop == "stop" & step >= (change_point - steps_to_include -1) & step < change_point-1 ~ "Late Adaptation",
    start_stop == "stop" & step > change_point & step < (change_point + steps_to_include +1) ~ "Early Post Adaptation",
    TRUE ~ NA_character_
  ),
    period = factor(period, levels = c("Late Baseline", "Early Adaptation", "Late Adaptation", "Early Post Adaptation"))) 


# Add the change points to the original plot
steps_by_period %>%
  ggplot(aes(x = step, y = max_speed_diff, color=period)) +
  geom_point() +
  geom_vline(data = speed_change_steps, aes(xintercept = change_point), color = "red", linetype = "dashed") +
  facet_grid(rows = vars(condition), cols = vars(start_stop)) +
  labs(title = "Belt Speed Difference with Detected Change Points",
       x = "Step",
       y = "Max Speed Difference")+
  scale_color_manual(values = c("Late Baseline" = "purple",
                              "Early Adaptation" = "red", 
                              "Late Adaptation" = "green", 
                              "Early Post Adaptation" = "blue",
                              "NA" = "grey"),  # Add this line
                   breaks = c("Late Baseline", "Early Adaptation", 
                              "Late Adaptation", "Early Post Adaptation"),
                   na.value = "grey") +  # Add this line
  theme(legend.position = "bottom") 

```

# Filter tidy_data

Get rid of any data that is related to steps that are not assigned to one of the periods of interest.

```{r}

tidy_data_by_period <- tidy_data %>% 
  left_join(steps_by_period) %>% 
  filter(!is.na(period)) %>% 
  select(-max_speed_diff, -change_point)

```
# Plot

```{r}

tidy_data_by_period %>% 
  filter(variable == "IPSI_ANKLE_MOMENT") %>% 
  filter(axis == "X") %>% 
  ggplot(aes(x = normalized_time,y=value, group=step))+
  geom_path()+
  facet_grid(cols = vars(period), rows = vars(condition))

tidy_data_by_period %>% 
  group_by(subject, order, condition, start_stop,stance_side,variable,axis,period, normalized_time) %>% 
  summarise(value = mean(value)) %>% 
  filter(variable == "IPSI_ANKLE_MOMENT") %>% 
  filter(axis == "X") %>% 
  ggplot(aes(x = normalized_time,y=value))+
  geom_path()+
  facet_grid(cols = vars(period), rows = vars(condition))

```


```{r}

tidy_data_by_period %>% 
  group_by(subject, order, condition, start_stop,stance_side,variable,axis,period, normalized_time) %>% 
  summarise(value = mean(value)) %>% 
  filter(variable %in% c("IPSI_ANKLE_MOMENT", "IPSI_ANKLE_ANGLE")) %>% 
  filter(period %in% c("Late Baseline", "Early Post Adaptation")) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  mutate(IPSI_ANKLE_MOMENT = -IPSI_ANKLE_MOMENT) %>% 
  filter(axis == "X") %>% 
  ggplot(aes(x = IPSI_ANKLE_ANGLE,y=IPSI_ANKLE_MOMENT, color=period))+
  geom_path()+
  facet_grid(cols = vars(condition))+
  theme(aspect.ratio = 1,
        legend.position = "bottom")
  

```


