---
title: "Ankle Quasi-stiffness"
author: "Mac Prible"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
options(error = function() traceback(2))
```


# Import Data
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)  # for p-values in mixed models
library(emmeans)
library(here)

dissertation_root <- normalizePath(file.path(here(),".."))
tidy_data_directory <- file.path(dissertation_root, "v3d", "tidy_output")
figure_directory <- file.path(dissertation_root, "figures")

mean_across_stance <- 
  list.files(tidy_data_directory, pattern="*_mean_across_stance.csv") %>% 
  map_df(~read_csv2(file.path(tidy_data_directory,.)))
  
```


## Define Conditions and Periods

```{r}
UNI_CONDITION_LABELS <- c(
  "sbt" = "Conventional SBT",
  "ubp" = "Uni. Fast Brake",
  "upp" = "Uni. Fast Prop"
)

BIL_CONDITION_LABELS <- c(
  "bbp" = "Bil. Fast Brake",
  "bpp" = "Bil. Fast Prop"
)

CONDITION_LABELS <- c(UNI_CONDITION_LABELS, BIL_CONDITION_LABELS)

all_periods <- mean_across_stance %>% distinct(period)
all_periods
pre_post_periods <- c("Late Baseline", "Early Post Adapt")
# pre_post_periods <- c("Late Baseline", "end stop")

```


# Stiffness Data
## Creation
```{r}

stiffness_data <- mean_across_stance %>% 
  filter(axis == "X") %>%  # interested in DF
  select(-axis) %>% 
  filter(period %in% pre_post_periods) %>%  # reduce the size of the data to manage
  mutate(period = factor(period, levels=c("Late Baseline", "Early Post Adapt"))) %>% 
  filter(variable %in% c("IPSI_ANKLE_ANGLE", "IPSI_ANKLE_MOMENT")) %>% 
  pivot_wider(names_from = variable,values_from = value) %>% 
  mutate(IPSI_ANKLE_MOMENT = -IPSI_ANKLE_MOMENT)

```

## Individual Quasi-Stiffness Curves
```{r fig.width=6, fig.height=10}

stiffness_data %>% 
  filter(subject == "S10") %>%
  ggplot(aes(x = IPSI_ANKLE_ANGLE, y = IPSI_ANKLE_MOMENT, color = period))+
  geom_path()+
  facet_grid(rows = vars(condition), cols = vars(stance_side))

```

# Key Event Identification

Using 3 phases from Molitor & Neptune (2024) Table 1: 

1. Dorsiflexion
2. Dual Flexion
3. Falling

## Point 1: Entering Dorsiflexion

First max plantarflexion angle. Note that given the definition of things here, this is calculated as the min. 

### Calculate Point

```{r}
pt_1_entering_dorsiflexion <- stiffness_data %>% 
  group_by(subject, condition, period, stance_side) %>%
  filter(normalized_time<50) %>% 
  slice(which.min(IPSI_ANKLE_ANGLE)) %>% 
  rename(max_PF_angle_time = normalized_time,
         max_PF_angle = IPSI_ANKLE_ANGLE) %>% 
  mutate(pt_1_entering_dorsiflexion_time = max_PF_angle_time)
```

### Visually Check Results
```{r, fig.width=6, fig.height=30}

stiffness_data %>% 
  # filter(subject == sample_subject) %>% 
  ggplot(aes(x= normalized_time,y = IPSI_ANKLE_ANGLE))+
  geom_path()+
  geom_point(data = pt_1_entering_dorsiflexion,
             aes(x =max_PF_angle_time, y = max_PF_angle),
             color="green")+
  facet_grid(rows = vars(subject, period), cols = vars(condition, stance_side)) 


```


## Point 2: Enter Dual Phase

### Reformat GRF Data 

Phase defined by APGRF and VGRF. Isolate these variables and rename for clarity. 

```{r}
grf_data <- mean_across_stance %>% 
  filter(period %in% pre_post_periods) %>% 
  filter(variable == "IPSI_GRF") %>% 
  mutate(period = factor(period, levels=c("Late Baseline", "Early Post Adapt"))) %>% 
  pivot_wider(names_from = axis, values_from = value) %>% 
  # select(-X) %>% 
  rename(VGRF = Z,
         APGRF = Y) 
```



### Find Critical Points and Average
```{r}
find_local_min_VGRF_time <- function(data) {
  # requires data be grouped: group_by(subject, condition, period, stance_side)
  # Find peak in first half
  first_peak <- data %>%
    filter(normalized_time <= 50) %>%
    slice(which.max(VGRF))
  
  # Find peak in second half
  second_peak <- data %>%
    filter(normalized_time > 50) %>%
    slice(which.max(VGRF))
  
  # Find minimum between the two peaks
  central_min <- data %>%
    filter(normalized_time > first_peak$normalized_time,
           normalized_time < second_peak$normalized_time) %>%
    slice(which.min(VGRF)) %>%
    select(normalized_time, VGRF) %>% 
    rename(local_min_VGRF_time=normalized_time,
           local_min_VGRF = VGRF)
  
  return(central_min)
}


find_APGRF_crossover_time <- function(data) {
  # requires data be grouped: group_by(subject, condition, period, stance_side)
  
  # look in the middle range, find where the absolute value "bounces" off of zero (i.e. the minimum)
  crossover <- data %>%    
  filter(normalized_time >= 25, normalized_time <= 75) %>%
  slice(which.min(abs(APGRF))) %>% 
  select(normalized_time, APGRF) %>% 
  rename(APGRF_crossover_time = normalized_time,
         APGRF_crossover = APGRF)
  
  
  return(crossover)
 
} 

pt_2_enter_dual_phase <- grf_data %>% 
  group_by(subject,order, condition, period, stance_side) %>%
  nest() %>%
  mutate(local_min_VGRF_time = map(data, find_local_min_VGRF_time),
         APGRF_crossover_time = map(data, find_APGRF_crossover_time)) %>%
  select(-data) %>% 
  unnest(cols = c(local_min_VGRF_time, APGRF_crossover_time)) %>% 
  mutate(pt_2_enter_dual_phase_time = (APGRF_crossover_time + local_min_VGRF_time)/2) %>% 
  ungroup()

```


### Visually Check Results

In plots below it appears that S9 lacks a "double hump" in VGRF for one of the trials. The code here effectively just uses 50% for that point, which I think is somewhat reasonable since it's then just getting averaged in with the APGRF crossover point. Might also be reasonable to throw out any measurements without a typical double hump.
```{r, fig.width=6, fig.height=30}

# VGRF plot
grf_data  %>% 
  # filter(subject==sample_subject) %>%
  ggplot() +
  facet_grid(rows = vars(subject, period), cols = vars(condition, stance_side)) +
  geom_line(aes(x = normalized_time, y = VGRF)) +
  geom_point(data = pt_2_enter_dual_phase,
             aes(x = local_min_VGRF_time, y = local_min_VGRF), 
             color = "red", size = 2)

# APGRF plot
grf_data %>% 
  # filter(subject==sample_subject) %>%
  ggplot() +
  facet_grid(rows = vars(subject, period), cols = vars(condition, stance_side)) +
  geom_line(aes(x = normalized_time, y = APGRF)) +
  geom_point(data = pt_2_enter_dual_phase,
             aes(x = APGRF_crossover_time, y = APGRF_crossover), 
             color = "blue", size = 2)


```


## Point 3: Enter Falling

Midpoint of peak DF angle and moment. Keep in mind that this must happen AFTER point 2, so that event time can be used to narrow the window of possibility.

### Find Critical Points and Average

```{r}

peak_df_angle <- stiffness_data %>% 
  group_by(subject, condition, period, stance_side) %>%
  left_join(pt_2_enter_dual_phase) %>% 
  filter(normalized_time>pt_2_enter_dual_phase_time) %>% 
  slice(which.max(IPSI_ANKLE_ANGLE)) %>% 
  rename(peak_df_angle_time = normalized_time,
         peak_df_angle = IPSI_ANKLE_ANGLE)

peak_df_moment <- stiffness_data %>% 
  group_by(subject, condition, period, stance_side) %>%
  left_join(pt_2_enter_dual_phase) %>% 
  filter(normalized_time>pt_2_enter_dual_phase_time) %>% 
  slice(which.max(IPSI_ANKLE_MOMENT)) %>% 
  rename(peak_df_moment_time = normalized_time,
         peak_df_moment = IPSI_ANKLE_MOMENT)


pt_3_enter_falling <- peak_df_angle %>% 
  left_join(peak_df_moment) %>% 
  select(-starts_with("pt_")) %>%  # don't want part 2 cluttering up
  mutate(pt_3_enter_falling_time = (peak_df_angle_time + peak_df_moment_time)/2)
  
```


### Visually Check Results
```{r, fig.width=6, fig.height=30}
  
stiffness_data %>% 
  ggplot(aes(x= normalized_time,y = IPSI_ANKLE_ANGLE))+
  geom_path()+
  geom_point(data = peak_df_angle,
             aes(x =peak_df_angle_time, y = peak_df_angle),
             color="blue")+
  facet_grid(rows = vars(subject, period), cols = vars(condition, stance_side)) 

stiffness_data %>% 
  ggplot(aes(x= normalized_time,y = IPSI_ANKLE_MOMENT))+
  geom_path()+
  geom_point(data = peak_df_moment,
             aes(x =peak_df_moment_time, y = peak_df_moment),
             color="red")+
  facet_grid(rows = vars(subject, period), cols = vars(condition, stance_side)) 
  
```

Note that Pt 4 is toe off, which is just 100%.


# Merge Key Points into single Dataframe

```{r}
  
all_key_points <- pt_1_entering_dorsiflexion %>% 
  left_join(pt_2_enter_dual_phase, by = join_by(subject, order, condition, period, stance_side)) %>%
  left_join(pt_3_enter_falling, by = join_by(subject, order, condition, period, stance_side)) %>% 
  select(subject, 
         order, 
         condition, 
         period, 
         stance_side,
         starts_with("pt_"))

```

# Assign Phase Based on Key Points


```{r}

stiffness_with_phase <- stiffness_data %>% 
  ungroup() %>% 
  left_join(all_key_points) %>% 
  mutate(Phase = case_when(
    normalized_time < pt_1_entering_dorsiflexion_time ~ "Initial",
    normalized_time >= pt_1_entering_dorsiflexion_time & normalized_time < pt_2_enter_dual_phase_time ~ "Dorsiflexion",
    normalized_time >= pt_2_enter_dual_phase_time & normalized_time < pt_3_enter_falling_time ~ "Dual Flexion",
    normalized_time >= pt_3_enter_falling_time ~ "Falling",
    TRUE ~ "None"  # catch any other cases
  )) %>% 
  mutate(Phase = factor(Phase, levels = c("Initial", "Dorsiflexion", "Dual Flexion", "Falling", "None")))
  
  
```

# Calculate All Quasi-stiffnesses
```{r}
# Update phase_stiffness to include intercepts
phase_stiffness <- stiffness_with_phase %>%
  filter(Phase != "Initial") %>% 
  group_by(subject, condition, period, stance_side, Phase) %>%
  summarise(
    stiffness = coef(lm(IPSI_ANKLE_MOMENT ~ IPSI_ANKLE_ANGLE))[2],  # slope
    intercept = coef(lm(IPSI_ANKLE_MOMENT ~ IPSI_ANKLE_ANGLE))[1],   # intercept
    r_squared = summary(lm(IPSI_ANKLE_MOMENT ~ IPSI_ANKLE_ANGLE))$r.squared
  ) %>%
  ungroup()


```

# Visually Check Quasi-stiffness Test Cases
```{r}
# Update the plotting code to use the calculated intercepts
phase_endpoints <- stiffness_with_phase %>%
  group_by(subject, condition, period, stance_side, Phase) %>%
  summarize(
    x_start = min(IPSI_ANKLE_ANGLE),
    x_end = max(IPSI_ANKLE_ANGLE)
  )

phase_lines <- phase_endpoints %>%
  left_join(phase_stiffness)

inspect_subject <- "S2"

# Plot with proper intercepts
stiffness_with_phase %>%
  filter(subject == inspect_subject) %>% 
  ggplot(aes(x = IPSI_ANKLE_ANGLE, y = IPSI_ANKLE_MOMENT)) +
  geom_path(aes(color = Phase)) +
  geom_segment(data = phase_lines %>% filter(subject == inspect_subject),
               aes(x = x_start, 
                   xend = x_end,
                   y = intercept + stiffness * x_start,
                   yend = intercept + stiffness * x_end,
                   color = Phase),
               linetype = "dashed",
               size = 1) +
  facet_grid(rows = vars(condition), cols = vars(stance_side, period)) +
  theme_minimal()
```

# Does the symmetrical PDSV alter ankle stiffness?
```{r}

phase_stiffness_for_test <- phase_stiffness %>% 
  # filter(stance_side == "left") %>%
  filter(subject != "S8") %>% # right ankle angle seems weird even in late baseline;
  # filter(subject != "S4") %>% 
  ungroup() %>% 
  filter(condition %in% c("bbp", "bpp")) %>%
  group_by(subject, condition, period, Phase) %>% 
  summarize(stiffness = mean(stiffness)) # average across both sides to reduce random noise.

phase_stiffness_for_test %>%
  group_by(subject, condition, period, Phase) %>% 
  summarize(stiffness = mean(stiffness)) %>% 
  ggplot(aes(x = Phase, y = stiffness, color = period)) +
  geom_point(position = position_dodge(width = 0.3), size = 1) +
  facet_grid(rows = vars(condition)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Stiffness (k)", 
       color = "Period")

```

```{r}

# Run separate models for each Phase since they represent different biomechanical events
for(current_phase in unique(phase_stiffness_for_test$Phase)) {
  
  cat("\nAnalysis for", current_phase, "phase:\n")
  
  # Subset data for current phase
  phase_data <- phase_stiffness_for_test %>%
    filter(Phase == current_phase,
           condition %in% c("bbp", "bpp"))
  
  # Fit mixed model
  model <- lmer(stiffness ~ period * condition + (1|subject), 
                data = phase_data)
  
  # Print ANOVA results
  print(anova(model))
}
```

```{r}


# For Dorsiflexion phase
dorsi_data <- phase_stiffness_for_test %>%
  filter(Phase == "Dorsiflexion",
         condition %in% c("bbp", "bpp"))
dorsi_model <- lmer(stiffness ~ period * condition + (1|subject), data = dorsi_data)

cat("\nDorsiflexion Phase - Within Condition Comparisons:\n")
dorsi_emm <- emmeans(dorsi_model, ~ period | condition)  # Note the | operator
pairs(dorsi_emm, adjust = "bonf")

# For Dual Flexion Phase
dual_flexion_data <- phase_stiffness_for_test %>%
  filter(Phase == "Dual Flexion",
         condition %in% c("bbp", "bpp"))
dual_flexion_model <- lmer(stiffness ~ period * condition + (1|subject), data = dual_flexion_data)
cat("\nDual Flexion Phase - Within Condition Comparisons:\n")
dual_flexion_emm <- emmeans(dual_flexion_model, ~ period | condition)  # Note the | operator
pairs(dual_flexion_emm, adjust = "bonf")

# For Falling phase
falling_data <- phase_stiffness_for_test %>%
  filter(Phase == "Falling",
         condition %in% c("bbp", "bpp"))
falling_model <- lmer(stiffness ~ period * condition + (1|subject), data = falling_data)

cat("\nFalling Phase - Within Condition Comparisons:\n")
falling_emm <- emmeans(falling_model, ~ period | condition)  # Note the | operator
pairs(falling_emm, adjust = "bonf")
```
