---
title: "Bilateral PDSV to Modulate Lower Extremity Joint Moments"
author: "Mac Prible"
output: html_document
---

```{r setup, include=FALSE}
library(here)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
options(error = function() traceback(2))

```

Color reference guide here: <https://sape.inf.usi.ch/quick-reference/ggplot2/colour>

# Import Data

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(cowplot) 
library(emmeans)
library(car)
library(clipr)
library(rstatix)
library(ggpubr)

theme_set(cowplot::theme_minimal_grid(font_family = "serif"))

tidy_data_directory <- here("data")
figure_directory <- here("manuscripts","figures")

raw_mean_across_stance <- 
  list.files(tidy_data_directory, pattern="*_mean_across_stance.csv") %>% 
  map_df(~read_csv2(file.path(tidy_data_directory,.)))

# note that step lengths here are handled separately because I don't want to average heel positions across multiple steps and then take the difference between them. Instead I take difference between heel positions at contact and then average across multiple steps. While they would likely generate similar results, I'm not 100% on that given the treadmill movement, so I took this more careful approach.  
# step_lengths <- 
#   list.files(tidy_data_directory, pattern="*_step_lengths.csv") %>% 
#   map_df(~read_csv2(file.path(tidy_data_directory,.)))

raw_mean_across_stance %>% 
  group_by(variable) %>% 
  summarize()

```

# Mild Data Adjustments

## Define Some Lists for Convenience

Some lists for convenience.

Periods:

1.  begin start\
2.  Late Baseline\
3.  Early Adapt\
4.  end start\
5.  begin stop\
6.  Late Adapt\
7.  Early Post Adapt
8.  end stop\

```{r}
periods <- c("begin start" = "begin start", 
             "Late Baseline" = "Late Baseline", 
             "Early Adapt" = "Early Adaptation",
             "end start" = "end start",
             "begin stop" = "begin stop",
             "Late Adapt" = "Late Adaptation",
             "Early Post Adapt" = "Early Post Adaptation",
             "end stop" = "end stop")

pre_post_periods <- c("Late Baseline", "Early Post Adaptation")
primary_periods <- c("Late Baseline", "Early Adaptation" , "Late Adaptation", "Early Post Adaptation")
pre_late_periods <- c("Late Baseline",  "Late Adaptation")
convergence_check <- c("begin stop", "Late Adaptation")

BIL_CONDITION_LABELS <- c(
  "bbp" = "Fast Brake",
  "bpp" = "Fast Prop"
)
```

## Assign Previous Condition

```{r}

previous_condition <- raw_mean_across_stance %>% 
  mutate(order = factor(order, levels = c(1:5))) %>% 
  group_by(subject, order, condition) %>% 
  summarize() %>% 
  ungroup() %>% 
  arrange(subject, order) %>%  
  group_by(subject) %>%
  mutate(previous_condition = lag(condition)) %>%
  mutate(previous_condition = replace_na(previous_condition, "none")) %>%
  ungroup() 

mean_across_stance <- raw_mean_across_stance %>% 
  ## Limit Data to Ipsilatearl Frame of reference, except for belt Speed
  filter(!str_detect(variable, "CONTRA") | str_detect(variable, "BeltSpeed") | str_detect(variable, "GRF")) %>% 
  ## Limit Data to Bilateral Conditions
  filter(condition %in% names(BIL_CONDITION_LABELS)) %>% 
  ##### More GENERAL FIXES ##############
  # make sure that order is a sensible factor 
  mutate(order = factor(order, levels = c(1:5))) %>% 
  # make sure that period is a sensible factor
  mutate(period = factor(period, 
                        levels = names(periods),
                        labels = unname(periods))) %>% 
  ## Average across sides
  group_by(subject, condition, period, normalized_time, axis, order, variable) %>% 
  summarize(value = mean(value)) %>% 
  ## Assign previous condition
  left_join(previous_condition) 

## Basic Checks You Have Data You Want
mean_across_stance %>% group_by(variable) %>% summarize()
mean_across_stance %>% group_by(condition) %>% summarize()
mean_across_stance %>% group_by(period) %>% summarize()
```

# Belt Speed

## Sample Participant

```{r}

measurement_names <- list("GRF" = "VGRF (N)",
                       "BeltSpeed" = "Belt Speed (m/s)")

# filter out the swing phase from contralateral belt speed
mean_across_stance %>% 
  filter(subject == "s1") %>%  
  filter(period == "Late Adaptation") %>% 
  filter(
    (variable %in% c("CONTRA_BeltSpeed", "IPSI_BeltSpeed") & axis == "X") |
    (variable %in% c("CONTRA_GRF", "IPSI_GRF") & axis == "Z")
  ) %>%
  select(-axis) %>% 
  ungroup() %>% 
  group_by(subject,  condition, period,variable, normalized_time) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = value, names_from = variable) %>% 
  mutate(CONTRA_BeltSpeed = case_when(CONTRA_GRF < 0.1 | normalized_time >= 30 & normalized_time <= 50 ~ NA_real_,
                                      TRUE ~ CONTRA_BeltSpeed),
         CONTRA_GRF = case_when(CONTRA_GRF < 0.1 | normalized_time >= 30 & normalized_time <= 50 ~ NA_real_,
                                      TRUE ~ CONTRA_GRF),
         ) %>% 
  # select(-contains("_GRF")) %>% 
  pivot_longer(cols = c(CONTRA_BeltSpeed, IPSI_BeltSpeed, CONTRA_GRF, IPSI_GRF)) %>% 
  mutate(measurement_type = case_when(str_detect(name, "GRF") ~ measurement_names$GRF,
                                      str_detect(name, "BeltSpeed") ~ measurement_names$BeltSpeed)) %>% 
  mutate(measurement_type = factor(measurement_type,levels=unlist(measurement_names))) %>% 
  mutate(name = case_when(str_detect(name, "IPSI") ~ "IPSI",
                          str_detect(name, "CONTRA") ~ "CONTRA",
                          TRUE ~ "")) %>% 
  mutate(name = factor(name, levels = c("IPSI", "CONTRA"))) %>% 
  ggplot(aes(x=normalized_time,y= value,color=name, group=subject))+
  geom_line(aes(group = interaction(subject, name)), size=0.5)+
  # scale_x_continuous(labels=NULL)+

  facet_grid(measurement_type~condition,
             scales="free_y",
             labeller = labeller(condition = BIL_CONDITION_LABELS))+
  theme(legend.position="bottom")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = NULL,
    title = "Example of Force Based Trigger for Belt Speed Changes",
    subtitle = "Average of last five steps of Late Adaptation for Participant 1",
    color = ""
  )+
  scale_color_manual(
    values = c("IPSI" = "blue", "CONTRA" = "red"),
    labels = c("IPSI" = "Reference Stance Limb", "CONTRA" = "Contralateral Limb") # Added labels for each color
  )
  
ggsave(file.path(figure_directory,"Example Force Based Trigger.svg"))

```

## all participants

```{r}

# filter out the swing phase from contralateral belt speed
mean_across_stance %>% 
  filter(period == "Late Adaptation") %>% 
  filter(
    (variable %in% c("CONTRA_BeltSpeed", "IPSI_BeltSpeed") & axis == "X") |
    (variable %in% c("CONTRA_GRF", "IPSI_GRF") & axis == "Z")
  ) %>%
  select(-axis) %>% 
  ungroup() %>% 
  group_by(subject,  condition, period,variable, normalized_time) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = value, names_from = variable) %>% 
  mutate(CONTRA_BeltSpeed = case_when(
    CONTRA_GRF < 0.1 | normalized_time >= 30 & normalized_time <= 50 ~ NA_real_,
    TRUE ~ CONTRA_BeltSpeed
  )) %>% 
  select(-contains("_GRF")) %>% 
  pivot_longer(cols = c(CONTRA_BeltSpeed, IPSI_BeltSpeed)) %>% 
  mutate(name = factor(name, levels = c("IPSI_BeltSpeed", "CONTRA_BeltSpeed"))) %>% 
  ggplot(aes(x=normalized_time,y= value,color=name, group=subject))+
  geom_line(aes(group = interaction(subject, name)), size=0.5)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(~condition, labeller = labeller(condition = BIL_CONDITION_LABELS))+
  cowplot::theme_minimal_grid(font_family="serif")+
  theme(legend.position="bottom",
        legend.justification = "center")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Belt Speed During Stance Phase",
    subtitle = "Average of Five Steps in Late Adaptation for Each Participant",
    color = ""
  )+
  scale_color_manual(
    values = c("IPSI_BeltSpeed" = "blue", "CONTRA_BeltSpeed" = "red"),
    labels = c("IPSI_BeltSpeed" = "Reference Limb", "CONTRA_BeltSpeed" = "Contralateral Limb") # Added labels for each color
  )
  
ggsave(file.path(figure_directory,"Bilateral PDSV speeds - All Participants.svg"))
```

## Fast Brake vs Fast Prop

```{r}
# First, create your filtered dataset
filtered_data <- mean_across_stance %>% 
  filter(axis == "X") %>% 
  filter(period == "Late Adaptation") %>% 
  filter(variable == "IPSI_BeltSpeed") %>% 
  group_by(subject, condition) %>% 
  summarize(Speed_Across_Stance = mean(value), .groups = "drop")

# Then calculate the summary statistics separately
summary_stats <- filtered_data %>%
  group_by(condition) %>%
  summarize(
    mean_val = mean(Speed_Across_Stance),
    sd_val = sd(Speed_Across_Stance),
    .groups = "drop"
  )

# Create the plot
ggplot(filtered_data, aes(x=condition, y=Speed_Across_Stance)) +
  geom_boxplot(width=.1) +
  geom_point() + 
  geom_line(aes(group = subject), alpha = 0.3) +
  stat_compare_means(method = "t.test", paired = TRUE, label.x.npc = "middle") +
  stat_summary(fun = mean, geom = "point", shape=23, fill= "red", color = "red", size = 2) +
  stat_summary(fun = mean, geom = "line", linetype = "dashed", size = 1.2, color = "red") +
  # Add the text annotations manually
  geom_text(
    data = summary_stats,
    aes(x = condition, y = mean_val, 
        label = sprintf("%.2f ± %.2f", mean_val, sd_val)),
    hjust = ifelse(summary_stats$condition == "bpp", -0.5, 1.5),
    vjust = 0.5,
    size = 4,
    color = "black"
  ) +
  labs(x = "Condition",
       title = "Mean Belt Speed Across Stance",
       y = "Speed (m/s)",
       subtitle = "")

```

While the mean belt speeds in the two conditions are statistically different, it amounts to a difference of .06 m/s. The


# Stance Times

Needed to de-normalize impulse with respect to time.

```{r}
## Get Stance Times
stance_times <- mean_across_stance %>% 
  filter(variable %in% c("TIME") & axis == "X") %>% 
  group_by(subject, condition, period, normalized_time) %>% 
  summarize(time = mean(value)) %>% 
  ungroup() %>% 
  group_by(subject, condition, period) %>%   
  summarize(start_stance = min(time),
            stop_stance = max(time)) %>% 
  ungroup() %>% 
  mutate(stance_time = stop_stance - start_stance)


stance_times %>% 
  ggplot(aes(x = period,y=stance_time, group = subject, color=condition))+
  geom_point()+
  facet_grid(~condition)
  
```

## Quality Check

### Condition Order or Previous Condition

```{r}
  
# Prepare data for lmer
bilateral_data$subject <- factor(bilateral_data$subject)
bilateral_data$condition <- factor(bilateral_data$condition)
bilateral_data$previous_condition <- factor(bilateral_data$previous_condition)
bilateral_data$period <- factor(bilateral_data$period)

model <- lmer(PF_impulse_adjusted ~ period * condition + order + previous_condition + (1|subject), 
             data = bilateral_data)


summary(model)

Anova(model, type = "III")
```

### Convergence and Stability in Late Adaptation

```{r}

PF_impulse %>% 
  ungroup() %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  ggplot(aes(x = period,y=PF_impulse_adjusted, group = subject, color=condition))+
  geom_boxplot(aes(group=period), 
               alpha=0.3, 
               width=0.2, 
               color="black") +
  geom_line(alpha=0.3, linewidth=0.3) +
  geom_point(alpha=0.5, size=1) + 
  facet_grid(~ condition) + 
  cowplot::theme_minimal_grid()+
  theme(
        # axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")+
  labs(
    x = "Period",
    y = "PF Impulse (N.m.s/BW)",
    title = "Plantarflexor Impulse Across Stance Phase by Period",
  ) +
  geom_text(stat = "summary", 
            fun = mean,
            aes(label = sprintf("%.2f", ..y..),
                vjust = -3.4, 
                # hjust = ifelse(period=="Late Baseline", 1.75, -.75),
                group = 1),
            color = "black", 
            size = 4)+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6)+ 
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) 


```

# Generate Primary Data for Study

```{r, fig.width=6, fig.height=8}

primary_data <- mean_across_stance %>% 
    filter(condition %in% names(BIL_CONDITION_LABELS)) %>%  
    filter(axis == "X") %>% 
    filter(variable %in% c("IPSI_ANKLE_MOMENT", 
                           "IPSI_KNEE_MOMENT", 
                           "IPSI_HIP_MOMENT",
                           "IPSI_ANKLE_ANGLE", 
                           "IPSI_KNEE_ANGLE", 
                           "IPSI_HIP_ANGLE")) %>% 
    mutate(joint = case_when(str_detect(variable, "HIP"    ) ~ "Hip",
                             str_detect(variable, "KNEE"  ) ~ "Knee",
                             str_detect(variable, "ANKLE") ~ "Ankle",
                             TRUE ~ ""),
           measurement_type= case_when(str_detect(variable,"MOMENT" ) ~ "Moment",
                                       str_detect(variable,"ANGLE"  ) ~ "Angle",
                                       TRUE ~ ""),
           variable = paste0(joint, "_", measurement_type)) %>% 
    select(-joint, -measurement_type) %>% 
    pivot_wider(names_from = variable, values_from = value) %>% 
    mutate(Hip_Moment = -Hip_Moment,  # align with typical "moment of support" style presentations
           Ankle_Moment = -Ankle_Moment, # align with typical "moment of support" style presentations
           Hip_Angle = -Hip_Angle,
           Ankle_Angle = -Ankle_Angle 
           ) %>%  
    pivot_longer(cols=contains(c("Hip", "Knee", "Ankle")), names_to = "variable") %>% 
    separate(col=variable,into=c("joint", "measurement_type"),sep="_") %>% 
    mutate(joint = factor(joint, levels=c("Hip","Knee", "Ankle"))) 
  
```

# Joint Kinetics and Kinematics Across stance

## Bootstrap Approach

### Calculate Bootstrap Bands Function

```{r}
# Function to calculate bootstrap simultaneous confidence bands for a SINGLE joint/condition/period combination
calculate_bootstrap_bands <- function(subset_data, confidence_level = 0.95, bootstrap_iterations = 400) {
  
  print("Starting bootstrap calculation...")
  print(paste("Data points:", nrow(subset_data)))
  
  # Get unique subjects and times
  subjects <- unique(subset_data$subject)
  times <- sort(unique(subset_data$normalized_time))
  n_subjects <- length(subjects)
  
  print(paste("Number of subjects:", n_subjects))
  print(paste("Number of time points:", length(times)))
  
  # Check if we have enough data
  if(n_subjects <= 1) {
    stop("Too few subjects (need at least 2)")
  }
  
  # Calculate the mean curve
  print("Calculating mean curve...")
  mean_curve <- subset_data %>%
    group_by(normalized_time) %>%
    summarize(
      mean_value = mean(value, na.rm = TRUE),
      original_se = sd(value, na.rm = TRUE)/sqrt(sum(!is.na(value))),
      n_obs = sum(!is.na(value)),
      .groups = 'drop'
    )
  
  # Store bootstrap resampled means
  print(paste("Performing bootstrap with", bootstrap_iterations, "iterations..."))
  bootstrap_means <- matrix(NA, nrow = length(times), ncol = bootstrap_iterations)
  
  # Perform bootstrap resampling
  set.seed(123)  # For reproducibility
                  
  for (b in 1:bootstrap_iterations) {
    if (b %% 10 == 0) print(paste("  Bootstrap iteration:", b))
    
    # Sample subjects WITH REPLACEMENT
    sampled_subjects <- sample(subjects, n_subjects, replace = TRUE)
    
    # Count occurrences of each subject (for first iteration)
    if (b == 1) {
      subject_counts <- table(sampled_subjects)
      print("  Subject sampling for first iteration:")
      print(subject_counts)
    }
    
    # Create resample dataset with proper subject repetition
    resample_data <- data.frame()
    
    for (subj in unique(sampled_subjects)) {
      subj_count <- sum(sampled_subjects == subj)
      if (subj_count > 0) {
        subj_data <- subset_data %>% filter(subject == subj)
        repeated_data <- subj_data[rep(1:nrow(subj_data), times = subj_count), ]
        resample_data <- bind_rows(resample_data, repeated_data)
      }
    }
    
    # Calculate mean for this bootstrap sample
    bs_means <- resample_data %>%
      group_by(normalized_time) %>%
      summarize(mean = mean(value, na.rm = TRUE), .groups = 'drop')
    
    # Store in our matrix
    idx <- match(bs_means$normalized_time, times)
    bootstrap_means[idx, b] <- bs_means$mean
  }
  
  # Check for NAs in bootstrap_means
  na_count <- sum(is.na(bootstrap_means))
  if(na_count > 0) {
    print(paste("Warning:", na_count, "NAs in bootstrap_means matrix"))
  }
  
  # Calculate bootstrap standard errors
  print("Calculating bootstrap standard errors...")
  bootstrap_se <- apply(bootstrap_means, 1, sd, na.rm = TRUE)
  
  # Calculate standardized deviations for each bootstrap sample
  print("Calculating standardized deviations...")
  max_devs <- numeric(bootstrap_iterations)
  for (b in 1:bootstrap_iterations) {
    standardized_devs <- abs(bootstrap_means[, b] - mean_curve$mean_value[match(times, mean_curve$normalized_time)]) / bootstrap_se
    max_devs[b] <- max(standardized_devs, na.rm = TRUE)
  }
  
  # Find critical value C that gives desired coverage
  critical_value <- quantile(max_devs, confidence_level, na.rm = TRUE)
  print(paste("Critical value for", confidence_level*100, "% confidence band:", critical_value))
  
  # Calculate t-critical for comparison
  t_critical <- qt(0.975, n_subjects-1)
  print(paste("t-critical value for comparison:", t_critical))
  
  # Calculate simultaneous confidence band
  print("Constructing confidence bands...")
  result <- mean_curve %>%
    rename(mean = mean_value) %>%
    mutate(
      se = bootstrap_se[match(normalized_time, times)],
      ci = critical_value * se,
      upper = mean + ci,
      lower = mean - ci,
      
      # Also calculate Gaussian for comparison
      gaussian_se = original_se,
      gaussian_ci = t_critical * gaussian_se,
      gaussian_upper = mean + gaussian_ci,
      gaussian_lower = mean - gaussian_ci,
      
      # Ratio for comparison
      ci_ratio = ci / gaussian_ci
    )
  
  return(result)
}


```


### Get Confidence Bands for All Conditions, Joints, Periods

```{r}

get_bootstrap_band_data <- function(moment_or_angle, bootstrap_iterations=400){
  
  # Filter to just the data we need
  filtered_data <- primary_data %>% 
    filter(period %in% pre_post_periods) %>% 
    filter(measurement_type == moment_or_angle)
  
  joints <- unique(filtered_data$joint)
  conditions <- unique(filtered_data$condition)
  periods <- unique(filtered_data$period)
  
  print("Unique values:")
  print(paste("Joints:", paste(joints, collapse=", ")))
  print(paste("Conditions:", paste(conditions, collapse=", ")))
  print(paste("Periods:", paste(periods, collapse=", ")))
  
  # Create all combinations manually
  combinations <- expand.grid(
    condition = conditions,
    joint = joints,
    period = periods,
    stringsAsFactors = FALSE
  )
  
  print(paste("Total combinations to process:", nrow(combinations)))
  
  # Now loop through these guaranteed unique combinations
  result_list <- list()
  
  for(i in 1:nrow(combinations)) {
    cond <- combinations$condition[i]
    jnt <- combinations$joint[i]
    per <- combinations$period[i]
    
    combo_name <- paste(cond, jnt, per, sep="_")
    print(paste("Processing combination", i, "of", nrow(combinations), ":", combo_name))
    
    # Get the subset of data for this combination
    subset_data <- filtered_data %>%
      filter(condition == cond, joint == jnt, period == per)
    
    # Only proceed if we have sufficient data
    if(nrow(subset_data) > 10) {  # Requiring at least 10 data points
      tryCatch({
        # Calculate bootstrap bands for this subset
        band_result <- calculate_bootstrap_bands(
          subset_data,
          confidence_level = 0.95,
          bootstrap_iterations = 400
        )
        
        # Add metadata
        band_result <- band_result %>%
          mutate(
            condition = cond,
            joint = jnt,
            period = per
          )
        
        # Store in results list
        result_list[[combo_name]] <- band_result
        
        print(paste("  Successfully processed:", combo_name))
      }, error = function(e) {
        print(paste("  ERROR processing:", combo_name, ":", e$message))
      })
    } else {
      print(paste("  Insufficient data for:", combo_name, "(", nrow(subset_data), "points )"))
    }
  }
  
  # Combine all data frames into a single one
  bootstrap_results <- bind_rows(result_list)
  
  bootstrap_results <- bootstrap_results %>%
    mutate(condition = BIL_CONDITION_LABELS[as.character(condition)])
  
  return(bootstrap_results)
}

```

### Plot Bootstrap Results

#### Moments
```{r,fig.width=6, fig.height=8}

moment_bootstrap_results <- get_bootstrap_band_data(moment_or_angle = "Moment", bootstrap_iterations = 400)

# Create the main bootstrap plot
moment_bootstrap_results %>% 
    ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
    geom_ribbon(aes(
      ymin = lower, 
      ymax = upper,
      fill = condition,
      group = interaction(condition, period)
    ), 
    alpha = 0.2,
    color = NA
    ) +
    geom_line(aes(y = mean), size = 1) +
    scale_x_continuous(breaks = seq(0, 100, by = 25)) +
    facet_grid(rows = vars(joint),
               cols = vars(condition),
               scales = "free_y") +
    scale_linetype_manual(values = setNames(c("dotted", "solid"), unique(moment_bootstrap_results$period))) +
    cowplot::theme_minimal_grid(font_family = "serif") +
    theme(
      legend.position = "bottom",
      legend.box.just = "center",
      legend.justification = "center"
    ) +
    labs(
      x = "Normalized Stance Time (0-100%)",
      y = "Joint Moment (N.m/BW)",
      title = "Joint Moment Across Stance (95% Confidence Bands)",
      subtitle = "Last 5 Steps of Baseline vs First 5 Steps of Post Adaptation",
      color = "Condition",
      fill = "Condition",
      linetype = "Period"
    )

ggsave(file.path(figure_directory,"pre_vs_post_moments_bootstrap.svg"), width = 6, height = 8)


```

#### Angles
```{r,fig.width=8, fig.height=8}

angle_bootstrap_results <- get_bootstrap_band_data(moment_or_angle = "Angle", bootstrap_iterations = 400)

# Create the main bootstrap plot
angle_bootstrap_results %>% 
    ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
    geom_ribbon(aes(
      ymin = lower, 
      ymax = upper,
      fill = condition,
      group = interaction(condition, period)
    ), 
    alpha = 0.2,
    color = NA
    ) +
    geom_line(aes(y = mean), size = 1) +
    scale_x_continuous(breaks = seq(0, 100, by = 25)) +
    facet_grid(rows = vars(joint),
               cols = vars(condition),
               scales = "free_y") +
    scale_linetype_manual(values = setNames(c("dotted", "solid"), unique(angle_bootstrap_results$period))) +
    cowplot::theme_minimal_grid(font_family = "serif") +
    theme(
      legend.position = "bottom",
      legend.box.just = "center",
      legend.justification = "center"
    ) +
    labs(
      x = "Normalized Stance Time (0-100%)",
      y = "Joint Angle (degrees)",
      title = "Joint Angle Across Stance (95% Confidence Bands)",
      subtitle = "Last 5 Steps of Baseline vs First 5 Steps of Post Adaptation",
      color = "Condition",
      fill = "Condition",
      linetype = "Period"
    )

ggsave(file.path(figure_directory,"pre_vs_post_angles_bootstrap.svg"), width = 6, height = 8)


```

### Plot all Individual data

This is just a sanity check of the bootstrapped results.

```{r}

primary_data %>% 
  filter(period %in% pre_post_periods) %>% 
  filter(joint=="Ankle") %>% 
  filter(axis=="X") %>% 
  filter(measurement_type=="Moment") %>%
  mutate(subject_period = paste0(subject, "_",period)) %>% 
  ggplot(aes(x=normalized_time,y=value, color = period, group=subject))+
  geom_line(aes(group=subject_period))+
  facet_grid(cols=vars(condition))
  
  

```

# Impulses

## Generate Impulse Data

```{r}
# comparison = pre_late_periods
comparison = pre_post_periods

moment_by_type <- primary_data %>% 
  filter(period %in% comparison) %>% 
  filter(measurement_type=="Moment") %>% 
  group_by(subject,joint, condition, period, normalized_time) %>%
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(moment_type = case_when(# Knee phases
                           joint == "Knee" & value > 0~ "Extension",
                           joint == "Knee" & value < 0~ "Flexion",
                           
                           # Ankle phases (based on moment sign)
                           joint == "Ankle" & value > 0 ~ "Extension", # or Plantarflexion
                           joint == "Ankle" & value < 0 ~ "Flexion", # or Dorsiflexion
                           
                           # Hip phases (based on moment sign)
                           joint == "Hip" & value > 0 ~ "Extension",
                           joint == "Hip" & value < 0 ~ "Flexion"
                           )
  ) 

impulse_data <- moment_by_type %>% 
  filter(moment_type %in% c("Extension", "Flexion")) %>% 
  group_by(subject, condition, period, joint, moment_type) %>% 
  summarize(impulse_normalized_time = abs(sum(value)/100)) %>% 
  left_join(stance_times) %>% 
  mutate(impulse_adjusted = impulse_normalized_time*stance_time)

```

## Test

```{r}

simple_comparison <- function(test_joint){
    print(paste0("Comparing changes for ", test_joint))
    impulse_data_for_test <- impulse_data%>% 
      filter(period %in% pre_post_periods) %>% 
      mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
      filter(joint==test_joint) %>% 
      filter(moment_type=="Extension")
    
    # Prepare data for lmer
    impulse_data_for_test$subject <- factor(impulse_data_for_test$subject)
    impulse_data_for_test$condition <- factor(impulse_data_for_test$condition)
    impulse_data_for_test$period <- factor(impulse_data_for_test$period)
    
    model <- lmer(impulse_adjusted ~ period * condition + (1|subject), 
                 data = impulse_data_for_test)
    
    
    print(summary(model))
    
    # Compute estimated marginal means
    emm <- emmeans(model, specs = ~ period | condition)
    
    # Pairwise comparisons of Period within each Condition
    period_comparisons <- pairs(emm, simple = "period", adjust="holm")
    summary(period_comparisons)
}

simple_comparison("Hip")
simple_comparison("Knee")
simple_comparison("Ankle")
```



## Plot

### Raw Plot

```{r, fig.height=8, fig.width=6}
moment_by_type %>% 
  ggplot(aes(x = normalized_time, y = value, color = condition, shape= period)) +
  geom_point(aes(group=subject))+
  geom_hline(yintercept = 0) +
  scale_x_continuous(labels = NULL) +
  facet_grid(cols = vars(period, condition), 
             rows = vars(joint, moment_type),
             scales = "free_y") +
  scale_linetype_manual(values = setNames(c("dotted", "solid"), comparison)) +
  cowplot::theme_minimal_grid() +
  theme(
    legend.position = "bottom",
    legend.box.just = "center",
    legend.justification = "center"
  ) +
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Joint Moment (N.m/BW)",
    title = "Average Joint Moment Across Stance (95% CI)",
    subtitle = "5 Steps Immediately Before and After Adaptation",
    color = NULL,
    fill = NULL,
    linetype=NULL
  )+
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")  



impulse_plot <- impulse_data %>% 
  filter(moment_type=="Extension") %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  ggplot(aes(x = period,y=impulse_adjusted, group = subject, color=condition))+
  geom_boxplot(aes(group=period), 
               alpha=0.3, 
               width=0.2, 
               color="black") +
  geom_line(alpha=0.3, linewidth=0.3) +
  geom_point(alpha=0.5, size=1) +
  facet_grid(joint~ condition,
             scales = "free_y")+
  cowplot::theme_minimal_grid(font_family = "serif")+
  theme(
        legend.position = "none")+
  # Add scale_x_discrete to create custom labels with line breaks
  scale_x_discrete(labels = c("Late\nBaseline", "Early Post\nAdaptation")) + 
  labs(
    x = "Period",
    y = "Joint Moment Impulse (N.m.s/BW)",
    title = "Angular Impulses Across Stance Phase by Period",
  ) +
  geom_text(stat = "summary",
            fun = mean,
            aes(label = sprintf("%.3f", ..y..),
                hjust = ifelse(period=="Late Baseline", 1.5, -.5),
                group = 1),
            color = "black",
            vjust = 0.4,
            size = 4)+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6)+ 
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) 

impulse_plot


```


### With Statistical Significance
```{r, fig.height=8, fig.width=6}
# Create a data frame with significance information
sig_data <- tibble(
  joint = factor(rep(c("Hip", "Knee", "Ankle"), each = 2), levels = c("Hip", "Knee", "Ankle")),
  condition = rep(c("Fast Brake", "Fast Prop"), 3),
  p_value = c("<0.0001", "<0.0001", "0.0048", "0.0048", "<0.0001", "<0.0001")
)

# Calculate appropriate y-positions for brackets based on your data
# You'll need to extract range information from your existing data
y_ranges <- impulse_data %>%
  filter(moment_type == "Extension") %>%
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>%
  group_by(joint, condition) %>%
  summarize(
    max_y = max(impulse_adjusted),
    min_y = min(impulse_adjusted),
    range = max_y - min_y,
    .groups = "drop"
  )

# Join with significance data and calculate positions
sig_data <- sig_data %>%
  left_join(y_ranges, by = c("joint", "condition")) %>%
  mutate(
    # Position brackets above the highest point with some padding
    y_pos = max_y + range * 0.10,
    # Scale bracket height based on the data range
    y_height = range * 0.05,
    # Convert p-values to significance symbols
    asterisks = case_when(
      p_value == "<0.0001" ~ "****",
      p_value == "0.0048" ~ "**",
      TRUE ~ ""
    )
  )

# Add significance indicators to your existing plot
impulse_plot_with_sig <- impulse_plot +
  # Add horizontal brackets
  geom_segment(data = sig_data,
               aes(x = 1, xend = 2, y = y_pos, yend = y_pos),
               inherit.aes = FALSE,
               linewidth = 0.5) +
  # Add vertical bracket ends
  geom_segment(data = sig_data,
               aes(x = 1, xend = 1, y = y_pos, yend = y_pos - y_height),
               inherit.aes = FALSE,
               linewidth = 0.5) +
  geom_segment(data = sig_data,
               aes(x = 2, xend = 2, y = y_pos, yend = y_pos - y_height),
               inherit.aes = FALSE,
               linewidth = 0.5) +
  # Add significance asterisks
  geom_text(data = sig_data,
            aes(x = 1.5, y = y_pos + y_height*0.7, label = asterisks),
            inherit.aes = FALSE,
            size = 3.5,
            family = "serif") +
  theme(plot.caption = element_text(hjust = 0, size = 9, family = "serif", face = "italic"))

# Display the enhanced plot
impulse_plot_with_sig

# Save to file if needed
ggsave(file.path(figure_directory, "Angular_Impulses_with_Significance.svg"), 
       impulse_plot_with_sig, width = 7, height = 8, dpi = 300)
```

### Tabular Form to aid written description

```{r}
impulse_data %>% 
  ungroup() %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  filter(moment_type=="Extension") %>% 
  group_by(joint, condition, period) %>% 
  summarize(
    mean = round(mean(impulse_adjusted), 3),
    se = round(sd(impulse_adjusted)/sqrt(n()), 3),   # Using standard error instead of SD
  ) %>% 
  mutate(period = str_replace(period, " ", "")) %>%
  pivot_wider(
    id_cols = c(joint, condition),
    names_from = period,
    values_from = c(mean, se),
    names_sep = "_"
  ) %>%
  # Calculate change and p-values from your model results
  mutate(
    Change = `mean_EarlyPost Adaptation` - mean_LateBaseline,
    # You would need to manually add p-values from your models
    pvalue = case_when(
      joint == "Hip" ~ "<0.0001",
      joint == "Knee" & condition == "Fast Brake" ~ "0.0048",
      joint == "Knee" & condition == "Fast Prop" ~ "0.0048",
      joint == "Ankle" & condition == "Fast Brake" ~ "<0.0001",
      joint == "Ankle" & condition == "Fast Prop" ~ "<0.0001",
      TRUE ~ "NA"
    )
  ) %>%
  # Format for display
  mutate(
    Baseline = paste0(mean_LateBaseline, " ± ", se_LateBaseline),
    PostAdaptation = paste0(`mean_EarlyPost Adaptation`, " ± ", `se_EarlyPost Adaptation`)
  ) %>%
  select(joint, condition, Baseline, PostAdaptation, Change, pvalue,`mean_LateBaseline`, `mean_EarlyPost Adaptation`)
```

### Correlation

Coming back to this I am realizing that this is not actually as big as I'd thought it might be. A future study using simulation could be conducted without this motiviation.

```{r, fig.width=5, fig.height=5}
impulse_data %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  filter(moment_type=="Extension") %>% 
  select(subject, condition, period, joint,impulse_adjusted) %>% 
  mutate(period = case_when(period=="Late Baseline" ~ "LB",
                            period=="Early Post Adaptation" ~ "EPA",
                            TRUE ~ "")) %>% 
  unite(measure_type, joint, period, sep="_") %>% 
  pivot_wider(names_from = measure_type, values_from = impulse_adjusted) %>% 
  mutate(delta_hip_ext_impulse = Hip_EPA - Hip_LB,
         delta_knee_ext_impulse = Knee_EPA - Knee_LB,
         delta_ankle_pf_impulse = Ankle_EPA - Ankle_LB) %>% 
  ggplot(aes(x = delta_ankle_pf_impulse, y = delta_knee_ext_impulse, color=condition))+
  geom_point()+ 
  stat_cor(aes(color=NULL),  # uses r.label instead of rr.label
           method = "pearson",
           p.accuracy = .001,
           label.x.npc = "middle", 
           label.y.npc = "top")+  
  stat_cor(aes(),  # uses r.label instead of rr.label
           method = "pearson",
           p.accuracy = .001,
           label.x.npc = "middle", 
           label.y.npc = "middle")+  
  theme(legend.position="bottom",
        legend.justification = "center",
        legend.title=element_blank())+
  labs(title="Change in Angular Impulse Before and After Adaptation",
       x="Δ Ankle Plantarflexor Impulse (N.m.s/BW)",
       y="Δ Knee Extensor Impulse (N.m.s/BW)")

  
```

