---
title: "Bilateral PDSV to Modulate Lower Extremity Joint Moments"
author: "Mac Prible"
output: html_document
---



```{r setup, include=FALSE}
library(here)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
options(error = function() traceback(2))

```

Color reference guide here: https://sape.inf.usi.ch/quick-reference/ggplot2/colour
  
# Import Data

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(cowplot) 
library(emmeans)
library(car)
library(clipr)
library(rstatix)
library(ggpubr)

theme_set(cowplot::theme_minimal_grid(font_family = "serif"))

tidy_data_directory <- here("data")
figure_directory <- here("manuscripts","figures")

raw_mean_across_stance <- 
  list.files(tidy_data_directory, pattern="*_mean_across_stance.csv") %>% 
  map_df(~read_csv2(file.path(tidy_data_directory,.)))

# note that step lengths here are handled separately because I don't want to average heel positions across multiple steps and then take the difference between them. Instead I take difference between heel positions at contact and then average across multiple steps. While they would likely generate similar results, I'm not 100% on that given the treadmill movement, so I took this more careful approach.  
# step_lengths <- 
#   list.files(tidy_data_directory, pattern="*_step_lengths.csv") %>% 
#   map_df(~read_csv2(file.path(tidy_data_directory,.)))

raw_mean_across_stance %>% 
  group_by(variable) %>% 
  summarize()

```


# Mild Data Adjustments

## Define Some Lists for Convenience

Some lists for convenience.

Periods:

1. begin start     
2. Late Baseline   
3. Early Adapt     
4. end start       
5. begin stop      
6. Late Adapt      
7. Early Post Adapt
8. end stop     
```{r}
periods <- c("begin start" = "begin start", 
             "Late Baseline" = "Late Baseline", 
             "Early Adapt" = "Early Adaptation",
             "end start" = "end start",
             "begin stop" = "begin stop",
             "Late Adapt" = "Late Adaptation",
             "Early Post Adapt" = "Early Post Adaptation",
             "end stop" = "end stop")

pre_post_periods <- c("Late Baseline", "Early Post Adaptation")
primary_periods <- c("Late Baseline", "Early Adaptation" , "Late Adaptation", "Early Post Adaptation")
pre_late_periods <- c("Late Baseline",  "Late Adaptation")
convergence_check <- c("begin stop", "Late Adaptation")

BIL_CONDITION_LABELS <- c(
  "bbp" = "Fast Brake",
  "bpp" = "Fast Prop"
)
```

## Assign Previous Condition
```{r}

previous_condition <- raw_mean_across_stance %>% 
  mutate(order = factor(order, levels = c(1:5))) %>% 
  group_by(subject, order, condition) %>% 
  summarize() %>% 
  ungroup() %>% 
  arrange(subject, order) %>%  
  group_by(subject) %>%
  mutate(previous_condition = lag(condition)) %>%
  mutate(previous_condition = replace_na(previous_condition, "none")) %>%
  ungroup() 

mean_across_stance <- raw_mean_across_stance %>% 
  ## Limit Data to Ipsilatearl Frame of reference, except for belt Speed
  filter(!str_detect(variable, "CONTRA") | str_detect(variable, "BeltSpeed") | str_detect(variable, "GRF")) %>% 
  ## Limit Data to Bilateral Conditions
  filter(condition %in% names(BIL_CONDITION_LABELS)) %>% 
  ##### More GENERAL FIXES ##############
  # make sure that order is a sensible factor 
  mutate(order = factor(order, levels = c(1:5))) %>% 
  # make sure that period is a sensible factor
  mutate(period = factor(period, 
                        levels = names(periods),
                        labels = unname(periods))) %>% 
  ## Average across sides
  group_by(subject, condition, period, normalized_time, axis, order, variable) %>% 
  summarize(value = mean(value)) %>% 
  ## Assign previous condition
  left_join(previous_condition) 

## Basic Checks You Have Data You Want
mean_across_stance %>% group_by(variable) %>% summarize()
mean_across_stance %>% group_by(condition) %>% summarize()
mean_across_stance %>% group_by(period) %>% summarize()
```

# Belt Speed

## Sample Participant

```{r}

measurement_names <- list("GRF" = "VGRF (N)",
                       "BeltSpeed" = "Belt Speed (m/s)")

# filter out the swing phase from contralateral belt speed
mean_across_stance %>% 
  filter(subject == "s1") %>%  
  filter(period == "Late Adaptation") %>% 
  filter(
    (variable %in% c("CONTRA_BeltSpeed", "IPSI_BeltSpeed") & axis == "X") |
    (variable %in% c("CONTRA_GRF", "IPSI_GRF") & axis == "Z")
  ) %>%
  select(-axis) %>% 
  ungroup() %>% 
  group_by(subject,  condition, period,variable, normalized_time) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = value, names_from = variable) %>% 
  mutate(CONTRA_BeltSpeed = case_when(CONTRA_GRF < 0.1 | normalized_time >= 30 & normalized_time <= 50 ~ NA_real_,
                                      TRUE ~ CONTRA_BeltSpeed),
         CONTRA_GRF = case_when(CONTRA_GRF < 0.1 | normalized_time >= 30 & normalized_time <= 50 ~ NA_real_,
                                      TRUE ~ CONTRA_GRF),
         ) %>% 
  # select(-contains("_GRF")) %>% 
  pivot_longer(cols = c(CONTRA_BeltSpeed, IPSI_BeltSpeed, CONTRA_GRF, IPSI_GRF)) %>% 
  mutate(measurement_type = case_when(str_detect(name, "GRF") ~ measurement_names$GRF,
                                      str_detect(name, "BeltSpeed") ~ measurement_names$BeltSpeed)) %>% 
  mutate(measurement_type = factor(measurement_type,levels=unlist(measurement_names))) %>% 
  mutate(name = case_when(str_detect(name, "IPSI") ~ "IPSI",
                          str_detect(name, "CONTRA") ~ "CONTRA",
                          TRUE ~ "")) %>% 
  mutate(name = factor(name, levels = c("IPSI", "CONTRA"))) %>% 
  ggplot(aes(x=normalized_time,y= value,color=name, group=subject))+
  geom_line(aes(group = interaction(subject, name)), size=0.5)+
  # scale_x_continuous(labels=NULL)+

  facet_grid(measurement_type~condition,
             scales="free_y",
             labeller = labeller(condition = BIL_CONDITION_LABELS))+
  theme(legend.position="bottom")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = NULL,
    title = "Example of Force Based Trigger for Belt Speed Changes",
    subtitle = "Average of last five steps of Late Adaptation for Participant 1",
    color = ""
  )+
  scale_color_manual(
    values = c("IPSI" = "blue", "CONTRA" = "red"),
    labels = c("IPSI" = "Reference Stance Limb", "CONTRA" = "Contralateral Limb") # Added labels for each color
  )
  
ggsave(file.path(figure_directory,"Example Force Based Trigger.svg"))

```

## all participants
```{r}

# filter out the swing phase from contralateral belt speed
mean_across_stance %>% 
  filter(period == "Late Adaptation") %>% 
  filter(
    (variable %in% c("CONTRA_BeltSpeed", "IPSI_BeltSpeed") & axis == "X") |
    (variable %in% c("CONTRA_GRF", "IPSI_GRF") & axis == "Z")
  ) %>%
  select(-axis) %>% 
  ungroup() %>% 
  group_by(subject,  condition, period,variable, normalized_time) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = value, names_from = variable) %>% 
  mutate(CONTRA_BeltSpeed = case_when(
    CONTRA_GRF < 0.1 | normalized_time >= 30 & normalized_time <= 50 ~ NA_real_,
    TRUE ~ CONTRA_BeltSpeed
  )) %>% 
  select(-contains("_GRF")) %>% 
  pivot_longer(cols = c(CONTRA_BeltSpeed, IPSI_BeltSpeed)) %>% 
  mutate(name = factor(name, levels = c("IPSI_BeltSpeed", "CONTRA_BeltSpeed"))) %>% 
  ggplot(aes(x=normalized_time,y= value,color=name, group=subject))+
  geom_line(aes(group = interaction(subject, name)), size=0.5)+
  # scale_x_continuous(labels=NULL)+
  facet_grid(~condition, labeller = labeller(condition = BIL_CONDITION_LABELS))+
  cowplot::theme_minimal_grid(font_family="serif")+
  theme(legend.position="bottom",
        legend.justification = "center")+
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Belt Speed (m/s)",
    title = "Belt Speed During Stance Phase",
    subtitle = "Average of Five Steps in Late Adaptation for Each Participant",
    color = ""
  )+
  scale_color_manual(
    values = c("IPSI_BeltSpeed" = "blue", "CONTRA_BeltSpeed" = "red"),
    labels = c("IPSI_BeltSpeed" = "Reference Limb", "CONTRA_BeltSpeed" = "Contralateral Limb") # Added labels for each color
  )
  
ggsave(file.path(figure_directory,"Bilateral PDSV speeds - All Participants.svg"))
```
## Fast Brake vs Fast Prop

```{r}


mean_across_stance %>% 
  filter(axis == "X") %>% 
  filter(period == "Late Adaptation") %>% 
  filter(variable == "IPSI_BeltSpeed") %>% 
  group_by(subject, condition) %>% 
  summarize(Speed_Across_Stance = mean(value)) %>%
  ggplot(aes(x=condition, y = Speed_Across_Stance))+
  geom_boxplot(width=.1) +
  geom_point() +
  geom_line(aes(group = subject), alpha = 0.3) +
  stat_compare_means(method = "t.test", paired = TRUE)+
  stat_summary(fun = mean, geom = "point",shape=23, fill= "red", color = "red", size = 2)+
  # Add dotted line between means
  stat_summary(fun = mean, geom = "line", linetype = "dashed", size = 1.2, color = "red") +
  # Add mean values as text
  stat_summary(fun = mean, 
              geom = "text", 
              aes(label = sprintf("%.2f", ..y..),
                  hjust = ifelse(condition == "bpp", -0.5, 1.5)), # shift left for bbp, right for bpp
              vjust = .5,
              size = 4,
              color = "black")+
  theme(title = "Mean Belt Speed",
        subtitle = )
  
```


# Characteristic Torque Profile

```{r}

# comparison <- pre_late_periods
comparison <- pre_post_periods

mean_across_stance %>% 
  filter(period %in% comparison) %>% 
  filter(condition %in% names(BIL_CONDITION_LABELS)) %>%  
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>%
  # filter(variable %in% c("IPSI_KNEE_MOMENT") & axis == "X") %>%
  group_by(period, condition, variable, normalized_time) %>% 
  summarize(
    PF_torque = mean(value),
    se = sd(value)/sqrt(n()),  # Standard error
    ci = 1.96 * se,            # 95% confidence interval
    upper = PF_torque + ci,
    lower = PF_torque - ci,
    .groups = 'drop'
  ) %>% 
  ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
  geom_ribbon(aes(
    ymin = lower, 
    ymax = upper,
    fill = condition,
    group = interaction(condition, period)
  ), 
  alpha = 0.2,
  color = NA
  ) +
  geom_line(aes(y = PF_torque), size = 1) +
  geom_hline(yintercept = 0) +
  scale_x_continuous(labels = NULL) +
  facet_grid(
    cols = vars(condition),
  ) +
  scale_linetype_manual(values = setNames(c("solid", "dashed"), comparison)) + 
  cowplot::theme_minimal_grid()+
  theme(legend.position = "bottom") +
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Ankle PF Torque (N.m/BW)",
    title = "Ankle Torque Across Conditions",
    subtitle = "Averaged Across Subjects (95% CI)",
    color = "Condition",
    fill = "Condition"
  )


ggsave(file.path(figure_directory,"torque_over_time.png"))
```

# PF Impulse

Needed to de-normalize impulse with respect to time.

```{r}
## Get Stance Times
stance_times <- mean_across_stance %>% 
  filter(variable %in% c("TIME") & axis == "X") %>% 
  group_by(subject, condition, period, normalized_time) %>% 
  summarize(time = mean(value)) %>% 
  ungroup() %>% 
  group_by(subject, condition, period) %>%   
  summarize(start_stance = min(time),
            stop_stance = max(time)) %>% 
  ungroup() %>% 
  mutate(stance_time = stop_stance - start_stance)

PF_impulse <- mean_across_stance %>% 
  # mutate(period = factor(period,levels = pre_post_periods)) %>% 
  filter(variable %in% c("IPSI_ANKLE_MOMENT") & axis == "X") %>% 
  mutate(PF_torque = value) %>% 
  select(-value, -axis, -variable) %>% 
  # group_by(subject, period, condition, order, variable, normalized_time,stance_side) %>% 
  # summarize(PF_torque = mean(value)) %>%
  filter(PF_torque<0) %>%
  group_by(subject, condition, previous_condition, order, period) %>% 
  summarize(PF_impulse_time_normalized = sum(PF_torque)/-100) %>% 
  left_join(stance_times) %>% 
  mutate(PF_impulse_adjusted = PF_impulse_time_normalized*stance_time)
  

```


```{r}
stance_times %>% 
  ggplot(aes(x = period,y=stance_time, group = subject, color=condition))+
  geom_point()+
  facet_grid(~condition)
  
```


## Plot
```{r}
plt <- PF_impulse %>% 
  filter(period %in% pre_post_periods) %>%
  ungroup() %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  group_by(subject,condition, period) %>% 
  summarize(PF_impulse= mean(PF_impulse_adjusted)) %>%  # averaging across sides
  # summarize(PF_impulse= mean(PF_impulse_time_normalized)) %>%  # averaging across sides
  ggplot(aes(x = period,y=PF_impulse, group = subject, color=condition))+
  geom_boxplot(aes(group=period), 
               alpha=0.3, 
               width=0.2, 
               color="black") +
  geom_line(alpha=0.3, linewidth=0.3) +
  geom_point(alpha=0.5, size=1) + 
  facet_grid(~ condition)+
  # scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  cowplot::theme_minimal_grid()+
  theme(
        # axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")+
  labs(
    x = "Period",
    y = "PF Impulse (N.m.s/BW)",
    title = "Plantarflexor Impulse Across Stance Phase by Period",
  ) +
  geom_text(stat = "summary", 
            fun = mean,
            aes(label = sprintf("%.2f", ..y..), 
                hjust = ifelse(period=="Late Baseline", 1.75, -.75),
                group = 1),
            color = "black", 
            vjust = 0.4, 
            size = 4)+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6)+ 
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) 

plt

ggsave(file.path(figure_directory,"bil_impulse_w_mean.svg"))
```

## Test

```{r}
bilateral_data <- PF_impulse %>% 
  ungroup() %>% 
  filter(period %in% pre_post_periods) %>%
  mutate(condition = BIL_CONDITION_LABELS[condition]) 
  
# Prepare data for lmer
bilateral_data$subject <- factor(bilateral_data$subject)
bilateral_data$condition <- factor(bilateral_data$condition)
bilateral_data$period <- factor(bilateral_data$period)

model <- lmer(PF_impulse_adjusted ~ period * condition + (1|subject), 
             data = bilateral_data)


summary(model)

# Compute estimated marginal means
emm <- emmeans(model, specs = ~ period | condition)

# Pairwise comparisons of Period within each Condition
period_comparisons <- pairs(emm, simple = "period", adjust="holm")
summary(period_comparisons)
```
## Quality Check

### Condition Order or Previous Condition

```{r}
  
# Prepare data for lmer
bilateral_data$subject <- factor(bilateral_data$subject)
bilateral_data$condition <- factor(bilateral_data$condition)
bilateral_data$previous_condition <- factor(bilateral_data$previous_condition)
bilateral_data$period <- factor(bilateral_data$period)

model <- lmer(PF_impulse_adjusted ~ period * condition + order + previous_condition + (1|subject), 
             data = bilateral_data)


summary(model)

Anova(model, type = "III")
```

### Convergence and Stability in Late Adaptation

```{r}

PF_impulse %>% 
  ungroup() %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  ggplot(aes(x = period,y=PF_impulse_adjusted, group = subject, color=condition))+
  geom_boxplot(aes(group=period), 
               alpha=0.3, 
               width=0.2, 
               color="black") +
  geom_line(alpha=0.3, linewidth=0.3) +
  geom_point(alpha=0.5, size=1) + 
  facet_grid(~ condition) + 
  cowplot::theme_minimal_grid()+
  theme(
        # axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")+
  labs(
    x = "Period",
    y = "PF Impulse (N.m.s/BW)",
    title = "Plantarflexor Impulse Across Stance Phase by Period",
  ) +
  geom_text(stat = "summary", 
            fun = mean,
            aes(label = sprintf("%.2f", ..y..),
                vjust = -3.4, 
                # hjust = ifelse(period=="Late Baseline", 1.75, -.75),
                group = 1),
            color = "black", 
            size = 4)+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6)+ 
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) 


```
# Generate Primary Data for Study

```{r, fig.width=6, fig.height=8}

primary_data <- mean_across_stance %>% 
    filter(condition %in% names(BIL_CONDITION_LABELS)) %>%  
    filter(axis == "X") %>% 
    filter(variable %in% c("IPSI_ANKLE_MOMENT", 
                           "IPSI_KNEE_MOMENT", 
                           "IPSI_HIP_MOMENT",
                           "IPSI_ANKLE_ANGLE", 
                           "IPSI_KNEE_ANGLE", 
                           "IPSI_HIP_ANGLE")) %>% 
    mutate(joint = case_when(str_detect(variable, "HIP"    ) ~ "Hip",
                             str_detect(variable, "KNEE"  ) ~ "Knee",
                             str_detect(variable, "ANKLE") ~ "Ankle",
                             TRUE ~ ""),
           measurement_type= case_when(str_detect(variable,"MOMENT" ) ~ "Moment",
                                       str_detect(variable,"ANGLE"  ) ~ "Angle",
                                       TRUE ~ ""),
           variable = paste0(joint, "_", measurement_type)) %>% 
    select(-joint, -measurement_type) %>% 
    pivot_wider(names_from = variable, values_from = value) %>% 
    mutate(Hip_Moment = -Hip_Moment,  # align with typical "moment of support" style presentations
           Ankle_Moment = -Ankle_Moment, # align with typical "moment of support" style presentations
           Hip_Angle = -Hip_Angle,
           Ankle_Angle = -Ankle_Angle 
           ) %>%  
    pivot_longer(cols=contains(c("Hip", "Knee", "Ankle")), names_to = "variable") %>% 
    separate(col=variable,into=c("joint", "measurement_type"),sep="_") %>% 
    mutate(joint = factor(joint, levels=c("Hip","Knee", "Ankle"))) 
  
```


# Baseline vs. Late Adaptation
## Moments
```{r, fig.height=8, fig.width=6}

comparison <-  pre_post_periods

primary_data %>% 
    filter(period %in% comparison) %>% 
    filter(measurement_type== "Moment") %>% 
    mutate(condition = BIL_CONDITION_LABELS[condition]) %>%   # more meaningful labels
    group_by(period, condition, joint, normalized_time) %>% 
    summarize(
      mean = mean(value),
      se = sd(value)/sqrt(n()),  # Standard error
      ci = 1.96 * se,            # 95% confidence interval
      upper = mean + ci,
      lower = mean - ci,
      .groups = 'drop'
    ) %>% 
    ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
    geom_ribbon(aes(
      ymin = lower, 
      ymax = upper,
      fill = condition,
      group = interaction(condition, period)
    ), 
    alpha = 0.2,
    color = NA
    ) +
    geom_line(aes(y = mean), size = 1) +
    scale_x_continuous(labels = NULL) +
    facet_grid(cols = vars(condition), 
               rows = vars(joint),
               scales = "free_y") +
    scale_linetype_manual(values = setNames(c("dotted", "solid"), comparison)) +
    cowplot::theme_minimal_grid(font_family = "serif") +
    theme(
      legend.position = "bottom",
      legend.box.just = "center",
      legend.justification = "center"
    ) +
    labs(
      x = "Normalized Stance Time (0-100%)",
      y = "Joint Moment (N.m/BW)",
      title = "Joint Moment Across Stance (95% CI)",
      subtitle = "Last 5 Steps of Baseline vs First 5 Steps of Post Adaptation",
      color = NULL,
      fill = NULL,
      linetype=NULL
    )+
    scale_color_discrete(guide = "none") +
    scale_fill_discrete(guide = "none")  


ggsave(file.path(figure_directory,"pre_vs_post_moments.svg"))
```

## Angles
```{r, fig.height=8, fig.width=6}


comparison <-  pre_post_periods

primary_data %>% 
    filter(period %in% comparison) %>% 
    filter(measurement_type== "Angle") %>% 
    mutate(condition = BIL_CONDITION_LABELS[condition]) %>%   # more meaningful labels
    group_by(period, condition, joint, normalized_time) %>% 
    summarize(
      mean = mean(value),
      se = sd(value)/sqrt(n()),  # Standard error
      ci = 1.96 * se,            # 95% confidence interval
      upper = mean + ci,
      lower = mean - ci,
      .groups = 'drop'
    ) %>% 
    ggplot(aes(x = normalized_time, color = condition, linetype = period)) +
    geom_ribbon(aes(
      ymin = lower, 
      ymax = upper,
      fill = condition,
      group = interaction(condition, period)
    ), 
    alpha = 0.2,
    color = NA
    ) +
    geom_line(aes(y = mean), size = 1) +
    scale_x_continuous(labels = NULL) +
    facet_grid(cols = vars(condition), 
               rows = vars(joint),
               scales = "free_y") +
    scale_linetype_manual(values = setNames(c("dotted", "solid"), comparison)) +
    cowplot::theme_minimal_grid(font_family = "serif") +
    theme(
      legend.position = "bottom",
      legend.box.just = "center",
      legend.justification = "center"
    ) +
    labs(
      x = "Normalized Stance Time (0-100%)",
      y = "Joint Angle (degrees)",
      title = "Joint Angle Across Stance (95% CI)",
      subtitle = "Last 5 Steps of Baseline vs First 5 Steps of Post Adaptation",
      color = NULL,
      fill = NULL,
      linetype=NULL
    )+
    scale_color_discrete(guide = "none") +
    scale_fill_discrete(guide = "none")  


ggsave(file.path(figure_directory,"pre_vs_post_angles.svg"))

```

## Impulses

### Plot

```{r, fig.height=8, fig.width=6}
# comparison = pre_late_periods
comparison = pre_post_periods

moment_by_type <- primary_data %>% 
  filter(period %in% comparison) %>% 
  filter(measurement_type=="Moment") %>% 
  group_by(subject,joint, condition, period, normalized_time) %>%
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(moment_type = case_when(# Knee phases
                           joint == "Knee" & value > 0~ "Extension",
                           joint == "Knee" & value < 0~ "Flexion",
                           
                           # Ankle phases (based on moment sign)
                           joint == "Ankle" & value > 0 ~ "Extension", # or Plantarflexion
                           joint == "Ankle" & value < 0 ~ "Flexion", # or Dorsiflexion
                           
                           # Hip phases (based on moment sign)
                           joint == "Hip" & value > 0 ~ "Extension",
                           joint == "Hip" & value < 0 ~ "Flexion"
                           )
  ) 

moment_by_type %>% 
  ggplot(aes(x = normalized_time, y = value, color = condition, shape= period)) +
  geom_point(aes(group=subject))+
  geom_hline(yintercept = 0) +
  scale_x_continuous(labels = NULL) +
  facet_grid(cols = vars(period, condition), 
             rows = vars(joint, moment_type),
             scales = "free_y") +
  scale_linetype_manual(values = setNames(c("dotted", "solid"), comparison)) +
  cowplot::theme_minimal_grid() +
  theme(
    legend.position = "bottom",
    legend.box.just = "center",
    legend.justification = "center"
  ) +
  labs(
    x = "Normalized Stance Time (0-100%)",
    y = "Joint Moment (N.m/BW)",
    title = "Average Joint Moment Across Stance (95% CI)",
    subtitle = "5 Steps Immediately Before and After Adaptation",
    color = NULL,
    fill = NULL,
    linetype=NULL
  )+
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none")  


impulse_data <- moment_by_type %>% 
  filter(moment_type %in% c("Extension", "Flexion")) %>% 
  group_by(subject, condition, period, joint, moment_type) %>% 
  summarize(impulse_normalized_time = abs(sum(value)/100)) %>% 
  left_join(stance_times) %>% 
  mutate(impulse_adjusted = impulse_normalized_time*stance_time)

impulse_data %>% 
  filter(moment_type=="Extension") %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  ggplot(aes(x = period,y=impulse_adjusted, group = subject, color=condition))+
  geom_boxplot(aes(group=period), 
               alpha=0.3, 
               width=0.2, 
               color="black") +
  geom_line(alpha=0.3, linewidth=0.3) +
  geom_point(alpha=0.5, size=1) +
  facet_grid(joint~ condition,
             scales = "free_y")+
  cowplot::theme_minimal_grid(font_family = "serif")+
  theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")+
  labs(
    x = "Period",
    y = "Joint Moment Impulse (N.m.s/BW)",
    title = "Angular Impulses Across Stance Phase by Period",
  ) +
  geom_text(stat = "summary",
            fun = mean,
            aes(label = sprintf("%.3f", ..y..),
                hjust = ifelse(period=="Late Baseline", 1.5, -.5),
                group = 1),
            color = "black",
            vjust = 0.4,
            size = 4)+
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "line", 
              color = "black", 
              linewidth = 1, 
              linetype = "dashed", 
              alpha = .6)+ 
  stat_summary(aes(group = 1), 
              fun = mean, 
              geom = "point", 
              color = "black", 
              size = 3, 
              shape = 18) 

ggsave(file.path(figure_directory,"Angular Impulses by Period.svg"))

```

### Tabular Form

```{r}

impulse_data %>% 
  ungroup() %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  filter(moment_type=="Extension") %>% 
  group_by(joint, moment_type,condition, period) %>% 
  summarize(
    mean = round(mean(impulse_adjusted),3),
    sd = round(sd(impulse_adjusted),3),
  ) %>% 
  mutate(period = str_replace(period, " ",""),
         condition = str_replace(condition, " ", ""),
         description = paste0(mean , "±", sd)) %>% 
  select(-mean, -sd) %>% 
  pivot_wider(names_from=period, values_from = description)



```

### Correlation

```{r, fig.width=7, fig.height=7}
impulse_data %>% 
  filter(joint %in% c("Knee", "Ankle")) %>% 
  mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
  filter(moment_type=="Extension") %>% 
  select(subject, condition, period, joint,impulse_adjusted) %>% 
  mutate(period = case_when(period=="Late Baseline" ~ "LB",
                            period=="Early Post Adaptation" ~ "EPA",
                            TRUE ~ "")) %>% 
  unite(measure_type, joint, period, sep="_") %>% 
  pivot_wider(names_from = measure_type, values_from = impulse_adjusted) %>% 
  mutate(delta_knee_ext_impulse = Knee_EPA - Knee_LB,
         delta_ankle_pf_impulse = Ankle_EPA - Ankle_LB) %>% 
  ggplot(aes(x = delta_ankle_pf_impulse, y = delta_knee_ext_impulse, color=condition))+
  geom_point()+ 
  # geom_text(aes(label = subject), 
  #         nudge_x = 0.01,    # adjust these values to offset the text
  #         nudge_y = 0.01) +  # from the points
  stat_cor(aes(color=NULL),  # uses r.label instead of rr.label
           method = "pearson",
           p.accuracy = .001,
           label.x.npc = "middle", 
           label.y.npc = "top")+  
  theme(legend.position="bottom",
        legend.justification = "center",
        legend.title=element_blank())+
  labs(title="Change in Angular Impulse Before and After Adaptation",
       x="Δ Ankle Plantarflexor Impulse (N.m.s/BW)",
       y="Δ Knee Extensor Impulse (N.m.s/BW)")

  
```


## Test

```{r}


simple_comparison <- function(test_joint){
    print(paste0("Comparing changes for ankle"))
    impulse_data_for_test <- impulse_data%>% 
      filter(period %in% pre_post_periods) %>% 
      mutate(condition = BIL_CONDITION_LABELS[condition]) %>% 
      filter(joint==test_joint) %>% 
      filter(moment_type=="Extension")
    
    # Prepare data for lmer
    impulse_data_for_test$subject <- factor(impulse_data_for_test$subject)
    impulse_data_for_test$condition <- factor(impulse_data_for_test$condition)
    impulse_data_for_test$period <- factor(impulse_data_for_test$period)
    
    model <- lmer(impulse_adjusted ~ period * condition + (1|subject), 
                 data = impulse_data_for_test)
    
    
    summary(model)
    
    # Compute estimated marginal means
    emm <- emmeans(model, specs = ~ period | condition)
    
    # Pairwise comparisons of Period within each Condition
    period_comparisons <- pairs(emm, simple = "period", adjust="holm")
    summary(period_comparisons)
}

simple_comparison("Hip")
simple_comparison("Knee")
simple_comparison("Ankle")
```

