---
title: "Quality Checks"
author: "Mac Prible"
output: html_document
---

This may turn into just a scratch pad to think through how to implement the checks that they committe wanted me to.

```{r setup, include=FALSE}
library(here)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
options(error = function() traceback(2))
```

 
 
# Import Data

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(cowplot) # Mac, Check out customization options
library(emmeans)

tidy_data_directory <- here("data")
figure_directory <- here("manuscripts","figures")

mean_across_stance <- 
  list.files(tidy_data_directory, pattern="*_mean_across_stance.csv") %>% 
  map_df(~read_csv2(file.path(tidy_data_directory,.)))
  
```



## Convenient Lists

Conditions: 

1. bbp      
2. bpp      
3. sbt      
4. ubp      
5. upp  

Periods:

1. begin start     
2. Late Baseline   
3. Early Adapt     
4. end start       
5. begin stop      
6. Late Adapt      
7. Early Post Adapt
8. end stop     


```{r}
UNI_CONDITION_LABELS <- c("sbt" = "Conventional SBT",
                          "ubp" = "Uni. Fast Brake",
                          "upp" = "Uni. Fast Prop")

BIL_CONDITION_LABELS <- c("bbp" = "Bil. Fast Brake",
                          "bpp" = "Bil. Fast Prop")

CONDITION_LABELS <- c(UNI_CONDITION_LABELS, BIL_CONDITION_LABELS)

pre_post_periods <- c("Late Baseline", 
                      "Early Post Adapt")

primary_periods <- c("Late Baseline", 
                     "Early Adapt",
                     "Late Adapt",
                     "Early Post Adapt")

pre_late_periods <- c("Late Baseline",
                      "Late Adapt")

ordered_periods <- c("Early Adapt",
                     "Early Post Adapt",
                     "Late Adapt",
                     "Late Baseline",
                     "begin start",
                     "begin stop",
                     "end start",
                     "end stop")

```

# Check SLR Convergence

To confirm that the adaption and wash out periods were of a sufficient length to allow both convergence to a stable pattern as well as a return to a comparable baseline, I'm going to examine the step length from multiple perspective. One basic one is the SLR for all of the asymmetrical conditions.

## Calculate step length across all conditions and start/stop of each period
```{r}

step_lengths <- mean_across_stance %>% 
  filter(normalized_time==1) %>% 
  filter(axis == "Y") %>% 
  filter(variable %in% c("IPSI_HEEL", "CONTRA_HEEL")) %>% 
  pivot_wider(names_from = variable,values_from = value) %>% 
  mutate(step_length = CONTRA_HEEL - IPSI_HEEL) %>% 
  mutate(period = factor(period, levels=ordered_periods)) %>% 
  mutate(order = factor(order, levels=c(1:5)))


slr <- step_lengths %>% 
  select(-CONTRA_HEEL, -IPSI_HEEL) %>%
  pivot_wider(names_from = stance_side,values_from = step_length) %>% 
  mutate(slr = left/right)
```

# Late Baseline Step Length 

```{r}

step_lengths %>% 
  filter(period %in% c("Late Baseline")) %>%
  ggplot(aes(x=condition, y = step_length))+
  geom_boxplot()+
  facet_grid(cols = vars(stance_side))+
  cowplot::theme_minimal_grid()

step_lengths %>% 
  filter(period %in% c("Late Baseline")) %>%
  ggplot(aes(x=order, y = step_length))+
  geom_boxplot()+
  facet_grid(cols = vars(stance_side))+
  cowplot::theme_minimal_grid()

```

# Step Length Ratio in Late Baseline

Is there any difference between 
```{r}

slr %>% 
  filter(condition %in% names(UNI_CONDITION_LABELS)) %>% 
  filter(period == "Late Baseline") %>% 
  ggplot(aes(x=condition,y=slr))+
  geom_boxplot()+
  geom_point()+
  cowplot::theme_minimal_grid()

```

